### 2021-01-25 Paper/Main.tex
paper update

\documentclass[12pt,oneside,reqno]{amsart}

\begin{document}
Hello World
\end{document}

### 2021-04-09 Code/stage_1_pre_rank.m
ESG

@@ -1,0 +1,71 @@
+% Matt Mike & Ben
+% Last Edited 2/24/21
+%% Setup Section
+% setting dataset for part 1 (please don't run unless you have
+% data_with_with_market_attached.csv)
+% MARET is market returns without dividendes (this while projected is w/o
+% div)
+clear; clc;
+A = readtable("data/ESG_and_five_factors.csv");
+%paremeters
+n_upper = 60; %max number of months
+n_lower = 12; %min number of months
+%%
+
+%individual stock var
+
+individual_stock_market_cap= A{:,10};
+stock_year_data = A{:,6};
+stock_month_data = A{:,7};
+stock_symbols_refrences = A{:,1};
+individual_stock_return = A{:,2};
+ESG_returns = A{:,3};
+risk_free = A{:,4};
+hml = A{:,5};
+smb = A{:,6};
+rmw = A{:,7};
+cma = A{:,8};
+cal_year = A{:,9};
+mmonth = A{:,10};
+
+market_reutrns = A{:,11};
+
+%% Actual Code 
+
+upper_counter =0;
+lower_counter = 0;
+pre_rank_matrix = nan(length(stock_symbols_refrences),12); %if I wanted individual beta we would extend this to 7
+for i =1:length(stock_symbols_refrences)-1
+    current_stock = stock_symbols_refrences(i);
+    next_stock =stock_symbols_refrences(i+1);
+    r_and_m_name = [stock_year_data(i)  stock_month_data(i) current_stock individual_stock_market_cap(i) individual_stock_return(i) risk_free(i) market_reutrns(i) cma(i) hml(i) smb(i) rmw(i)];%year and month and name
+    
+    if current_stock == next_stock
+        if lower_counter-1 >= n_lower
+            if upper_counter > n_upper
+                upper_counter = n_upper;
+            end
+            market = [market_reutrns(1+i-n_lower-upper_counter:i+1) market_reutrns(i-n_lower-upper_counter:i)];
+            individual = [individual_stock_return(1+i-n_lower-upper_counter:i+1)]; 
+            beta = regress(individual , market)';
+            sum_o_beta = sum(beta);
+            pre_rank_matrix(i,:)= [r_and_m_name sum_o_beta];
+            upper_counter = upper_counter + 1; 
+        else
+            pre_rank_matrix(i,:)= [r_and_m_name NaN];
+        end
+        
+        lower_counter = lower_counter+1;
+    else
[…]

### 2021-04-09 Paper/Main.tex
Updates from Overleaf

@@
 \begin{document}
-Hello World
+Over the last decade investors have become increasingly concerned with extra factors outside of financials to address the performance of stocks.
+Research from Mckendsey has shown that there is a direct correlation between diversity inside of executive team leadership and company performance.
+A rising threat of global climate related issues and workplace culture to determine top talent have all become extremely important for company management to consider on the basis of long term performance.
+
+this paper will be broken up into four major sections.
+We will first asses a brief literature review of papers looking abnormal performance around ESG criteria as well as discussion of possible models to apply asset pricing to ESG factors.
+Next I will discuss the data used in this papers as well as the data cleaning process and summary statistics.
+Following the Data will be a breakdown of the methodology for the paper.
+The final two sections will provide the results of the regressions and a conclusion and possible next steps.
+\section{Methodology}
+As dicussed in the literature review there are numerous different models for assessing asset pricing and abnormal returns.
+In this paper I will be using an extension on on Fama French's 2015 paper ``A five factor asset pricing model".
+
+To capture abnormal returns we will consider 5 factors that is the 5 factors from the famma french paper small minus big(SMB),CMA, HML, and market portfolio(I.e model, consisting of market, size, value, profitability, and investment factors;).
+In addition I will ad another term to address how stocks move with the movement to a ESG porlfio.
+For this purpose we will use the MSCI ESG all world index.
+
+\begin{equation}
+\begin{split}
+    R_{it}^e = a_i+b_i(R_{Mt}-R_{Fr})+s_iSMB_t+ & h_iHML_t+ & r_iRMW_t+c_iCMA+e_i(R_{ESG t}-R_{rf})
+    \end{split}
+\end{equation}
+where b
 \end{document}

### 2021-04-09 Paper/Main.tex
overleaf

--- 2021-01-25 Paper/Main.tex
@@
 \documentclass[12pt,oneside,reqno]{amsart}
+\input{Paper/settings}
 
 \begin{document}
-Hello World
+\section{Introduction and Motivation}
+Over the last decade investors have become increasingly concerned with extra factors outside of financials to address the performance of stocks.
+Research from Mckendsey has shown that there is a direct correlation between diversity inside of executive team leadership and company performance.
+A rising threat of global climate related issues and workplace culture to determine top talent have all become extremely important for company management to consider on the basis of long term performance.
+
+this paper will be broken up into five  sections.
+\textit{Section \ref{Literature}} will provide a  brief literature review of papers looking abnormal performance around ESG criteria as well as discussion of possible asset pricing models to capture ESG factors in performance.
+Next in  \textit{ Section \ref{Data}}  I will discuss the data used in this papers as well as the data cleaning process and summary statistics.
+Following the Data, \textit{Section \ref{Methodology}} will breakdown of the methodology for the paper.
+Finally, \textit{Section \ref{Results}}  will summarize the results of the regressions, \textit{Section \ref{Robustness}} will check the robustness of the results and \textit{Section \ref{Conclusion}} will provide a conclusion and possible extensions for next steps.
+
+\section{Literature Review}
+\label{Literature}
+
+\section{Data}
+\label{Data}
+The data used for this paper was soured three databases.
+The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices.
+Data on the ESG portfolio was extracted from compustat as the daily average returns for MSCI World ESG Focus Index.
+Finally, daily Fama French factors for SMB, CMA,HML, and RMW were sourced from the Kenneth R.
+French - Data Library.
+The time period for the data for the paper is between June 2011 and December 2020.
+The segment of stock data is determined by the length of available performance for the MSCI World ESG Focus Index as this index was initiated in June 2011.
+Daily values were summed and collapsed to created monthly values for factor and individual security performance.
+\section{Methodology}
+\label{Methodology}
+As dicussed in the literature review there are numerous different models for assessing asset pricing and abnormal returns.
+In this paper I will be using an extension on on Fama French's 2015 paper ``A five factor asset pricing model".
+
+To capture abnormal returns we will consider 5 factors that is the 5 factors from the famma french paper small minus big(SMB),CMA, HML, and market portfolio(I.e model, consisting of market, size, value, profitability, and investment factors;).
+In addition I will ad another term to address how stocks move with the movement to a ESG porlfio.
+For this purpose we will use the MSCI ESG all world index.
+
+\begin{equation}
+\begin{split}
+    R_{it}^e = a_i+b_i(R_{Mt}-R_{fr})+s_iSMB_t+ & h_iHML_t+ & r_iRMW_t+c_iCMA+e_i(R_{ESG t}-R_{rf})
+    \end{split}
+    \label{six_factor_basic_reg}
+\end{equation}
+In equation \eqref{six_factor_basic_reg} $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms, $e^e_{it}$ is the return of the individual secutiy minus the risk freee rate ($R_{it}-R_{rf}$
+
+\section{Results}
+\label{Results}
+
+\section{Robustness Checks}
+\label{Robustness}
+
+\section{Conclusion}
+\label{Conclusion}
+\section{Code}
+\textbf{Data Cleaning and setting up for Pre-Ranking}
+
+\lstinputlisting{Code/stage_1_pre_rank.m}
+\lstinputlisting[language=Stata]{Code/code_for_returns.do}
+
 \end{document}

### 2021-04-09 Paper/stata_code.tex
overleaf

\definecolor{codeblue}{rgb}{0.29296875, 0.51953125, 0.68359375}
\definecolor{codegreen}{rgb}{0.47265625, 0.62890625, 0.40234375}
\definecolor{codegray}{rgb}{0.95703125, 0.95703125, 0.95703125}
\definecolor{codecrimson}{rgb}{0.87109375,0.3984375,0.3984375}

\lstdefinelanguage{Stata}{
    % System commands
    morekeywords=[1]{regress, reg, summarize, sum, display, di, generate, gen, bysort, use, import, delimited, predict, quietly, probit, margins, test},
    % Reserved words
    morekeywords=[2]{aggregate, array, boolean, break, byte, case, catch, class, colvector, complex, const, continue, default, delegate, delete, do, double, else, eltypedef, end, enum, explicit, export, external, float, for, friend, function, global, goto, if, inline, int, local, long, mata, matrix, namespace, new, numeric, NULL, operator, orgtypedef, pointer, polymorphic, pragma, private, protected, public, quad, real, return, rowvector, scalar, short, signed, static, strL, string, struct, super, switch, template, this, throw, transmorphic, try, typedef, typename, union, unsigned, using, vector, version, virtual, void, volatile, while,},
    % Keywords
    morekeywords=[3]{forvalues, foreach, set},
    % Date and time functions
    morekeywords=[4]{bofd, Cdhms, Chms, Clock, clock, Cmdyhms, Cofc, cofC, Cofd, cofd, daily, date, day, dhms, dofb, dofC, dofc, dofh, dofm, dofq, dofw, dofy, dow, doy, halfyear, halfyearly, hh, hhC, hms, hofd, hours, mdy, mdyhms, minutes, mm, mmC, mofd, month, monthly, msofhours, msofminutes, msofseconds, qofd, quarter, quarterly, seconds, ss, ssC, tC, tc, td, th, tm, tq, tw, week, weekly, wofd, year, yearly, yh, ym, yofd, yq, yw,},
    % Mathematical functions
    morekeywords=[5]{abs, ceil, cloglog, comb, digamma, exp, expm1, floor, int, invcloglog, invlogit, ln, ln1m, ln, ln1p, ln, lnfactorial, lngamma, log, log10, log1m, log1p, logit, max, min, mod, reldif, round, sign, sqrt, sum, trigamma, trunc,},
    % Matrix functions
    morekeywords=[6]{cholesky, coleqnumb, colnfreeparms, colnumb, colsof, corr, det, diag, diag0cnt, el, get, hadamard, I, inv, invsym, issymmetric, J, matmissing, matuniform, mreldif, nullmat, roweqnumb, rownfreeparms, rownumb, rowsof, sweep, trace, vec, vecdiag, },
    % Programming functions
    morekeywords=[7]{autocode, byteorder, c, _caller, chop, abs, clip, cond, e, fileexists, fileread, filereaderror, filewrite, float, fmtwidth, has_eprop, inlist, inrange, irecode, matrix, maxbyte, maxdouble, maxfloat, maxint, maxlong, mi, minbyte, mindouble, minfloat, minint, minlong, missing, r, recode, replay, return, s, scalar, smallestdouble,},
    % Random-number functions
    morekeywords=[8]{rbeta, rbinomial, rcauchy, rchi2, rexponential, rgamma, rhypergeometric, rigaussian, rlaplace, rlogistic, rnbinomial, rnormal, rpoisson, rt, runiform, runiformint, rweibull, rweibullph,},
    % Selecting time-span functions
    morekeywords=[9]{tin, twithin,},
    % Statistical functions
    morekeywords=[10]{betaden, binomial, binomialp, binomialtail, binormal, cauchy, cauchyden, cauchytail, chi2, chi2den, chi2tail, dgammapda, dgammapdada, dgammapdadx, dgammapdx, dgammapdxdx, dunnettprob, exponential, exponentialden, exponentialtail, F, Fden, Ftail, gammaden, gammap, gammaptail, hypergeometric, hypergeometricp, ibeta, ibetatail, igaussian, igaussianden, igaussiantail, invbinomial, invbinomialtail, invcauchy, invcauchytail, invchi2, invchi2tail, invdunnettprob, invexponential, invexponentialtail, invF, invFtail, invgammap, invgammaptail, invibeta, invibetatail, invigaussian, invigaussiantail, invlaplace, invlaplacetail, invlogistic, invlogistictail, invnbinomial, invnbinomialtail, invnchi2, invnF, invnFtail, invnibeta, invnormal, invnt, invnttail, invpoisson, invpoissontail, invt, invttail, invtukeyprob, invweibull, invweibullph, invweibullphtail, invweibulltail, laplace, laplaceden, laplacetail, lncauchyden, lnigammaden, lnigaussianden, lniwishartden, lnlaplaceden, lnmvnormalden, lnnormal, lnnormalden, lnwishartden, logistic, logisticden, logistictail, nbetaden, nbinomial, nbinomialp, nbinomialtail, nchi2, nchi2den, nchi2tail, nF, nFden, nFtail, nibeta, normal, normalden, npnchi2, npnF, npnt, nt, ntden, nttail, poisson, poissonp, poissontail, t, tden, ttail, tukeyprob, weibull, weibullden, weibullph, weibullphden, weibullphtail, weibulltail,},
    % String functions 
    morekeywords=[11]{abbrev, char, collatorlocale, collatorversion, indexnot, plural, plural, real, regexm, regexr, regexs, soundex, soundex_nara, strcat, strdup, string, strofreal, string, strofreal, stritrim, strlen, strlower, strltrim, strmatch, strofreal, strofreal, strpos, strproper, strreverse, strrpos, strrtrim, strtoname, strtrim, strupper, subinstr, subinword, substr, tobytes, uchar, udstrlen, udsubstr, uisdigit, uisletter, ustrcompare, ustrcompareex, ustrfix, ustrfrom, ustrinvalidcnt, ustrleft, ustrlen, ustrlower, ustrltrim, ustrnormalize, ustrpos, ustrregexm, ustrregexra, ustrregexrf, ustrregexs, ustrreverse, ustrright, ustrrpos, ustrrtrim, ustrsortkey, ustrsortkeyex, ustrtitle, ustrto, ustrtohex, ustrtoname, ustrtrim, ustrunescape, ustrupper, ustrword, ustrwordcount, usubinstr, usubstr, word, wordbreaklocale, worcount,},
    % Trig functions
    morekeywords=[12]{acos, acosh, asin, asinh, atan, atanh, cos, cosh, sin, sinh, tan, tanh,},
    morecomment=[l]{//},
    % morecomment=[l]{*},  // `*` maybe used as multiply operator. So use `//` as line comment.
    morecomment=[s]{/*}{*/},
    % The following is used by macros, like `lags'.
    morestring=[b]{`}{'},
    % morestring=[d]{'},
    morestring=[b]",
    morestring=[d]",
    % morestring=[d]{\\`},
    % morestring=[b]{'},
    sensitive=true,
}

### 2021-04-09 Paper/Main.tex
Updates from Overleaf

@@
 \end{equation}
-In equation \eqref{six_factor_basic_reg} $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms, $e^e_{it}$ is the return of the individual secutiy minus the risk freee rate ($R_{it}-R_{rf}$
+In equation \eqref{six_factor_basic_reg} $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms, $e^e_{it}$ is the return of the individual secutiy minus the risk freee rate ($R_{it}-R_{rf}$)
 

### 2021-04-09 Code/stage_1_pre_rank.m
updating pre_rank2

@@ -1,5 +1,5 @@
-% Matt Mike & Ben
-% Last Edited 2/24/21
+% Matt 
+% Last Edited 4/9/21
 %% Setup Section
 % setting dataset for part 1 (please don't run unless you have
 % data_with_with_market_attached.csv)

### 2021-04-10 Code/stage_1_pre_rank.m
JIK

@@ -60,6 +60,6 @@ for i =1:length(stock_symbols_refrences)-1
     
 end 
 %%
-pre_rank_table = array2table(pre_rank_matrix,'VariableNames',{'annual','mmonth','kypermno','mtcap','mretx','risk_free','maret','cma','hml','smb','rmw','pre_ranked_beta'});%we changed myear to annual because of the porflio data uses annual
-writetable( pre_rank_table, 'pre_rank_table.csv')
+pre_rank_table = array2table(pre_rank_matrix,'VariableNames',{'kypermno','expected_return_stock','esg_minus_rf','market_minus_rf','hml','smb','rmw','cma','tcap','cal_year','mmonth','pre_ranked_beta_market','pre_ranked_esg'});%we changed myear to annual because of the porflio data uses annual
+writetable( pre_rank_table, 'data/pre_rank_table.csv')
 

### 2021-04-10 Code/cross_sectional_matlab.m
last step before famma mcbeth

@@ -1,0 +1,72 @@
+%Matt Chistolini
+%Last edited 3/5/21
+clear; clc;
+A = readtable("out_of_stata_five_factor.csv");
+porflio_number = A{:,3};
+mtrex = A{:,4};
+hml = A{:,5};
+smb = A{:,6};
+rmw = A{:,7}
+cma = A{:,8};
+modate = A{:,9};
+time_index = A{:,10};
+
+%% the show!!
+out_for_reg_1 = [];
+counter = 0;
+for i = 1:length(porflio_number)-1
+    
+    current_stock = porflio_number(i);
+    next_stock =porflio_number(i+1);
+    
+    if current_stock == 100;
+        returns_i = mtrex(i-counter:end); 
+        hml_i = hml(i-counter:end);
+        smb_i = smb(i-counter:end);
+        rmw_i= rmw(i-counter:end);
+        cma_i = cma(i-counter:end);
+        factors = [hml_i smb_i rmw_i cma_i ];
+        reg_1 = regress(returns_i, factors);
+        out_for_reg_1 = [out_for_reg_1 reg_1];
+        break
+    end
+    
+    if current_stock == next_stock
+        
+        counter = counter+1;
+    else
+        if counter >2
+        returns_i = mtrex(i-counter:i); 
+        hml_i = hml(i-counter:i);
+        smb_i = smb(i-counter:i);
+        rmw_i= rmw(i-counter:i);
+        cma_i = cma(i-counter:i);
+        factors = [hml_i smb_i rmw_i cma_i ];
+        reg_1 = regress(returns_i, factors);
+        out_for_reg_1 = [out_for_reg_1 reg_1];
+        counter = 0;
+        end 
+    end
+end
+
+%% Part 2: Second Regression
+
+time_length = max(time_index);
+out_for_reg_2 =[];
+counter = 0;
+for t = 1:1:time_length
+    counter=counter+1;
+    index_values = find(time_index ==t);
+    reg_2 = regress(mtrex(index_values),out_for_reg_1'); 
+    out_for_reg_2 = [out_for_reg_2 reg_2];
+end   
+%% finding averages and STD
+
+beta_avg = mean(out_for_reg_2')*100 % in percent 
+beta_std = std(out_for_reg_2')./sqrt(time_length)
+
+
+
+
+
+

### 2021-04-10 Code/stage_1_pre_rank.m

@@ -1,4 +1,4 @@
-% uthor Matt 
+% Author Matt Chistolini
 % Last Edited 4/9/21
 %% Setup Section
 clear; clc;

### 2021-04-10 Paper/Main.tex
Updates from Overleaf

@@
 Daily values were summed and collapsed to created monthly values for factor and individual security performance.
+
+we looking 830,891 observation for  12,194 unique stocks which gives an average of 68months of data
+
+to reduce some of the tail behavior we dropped the top and bottom  4\% of pre ranked beta values as well as companies with market values less tham 10 million
 \section{Methodology}
@@
 \section{Code}
-\textbf{Data Cleaning and setting up for Pre-Ranking}
-
-\lstinputlisting{Code/stage_1_pre_rank.m}
-\lstinputlisting[language=Stata]{Code/code_for_returns.do}
-
+\input{Paper/Code_files}
 \end{document}

### 2021-04-11 Paper/Main.tex
Updates from Overleaf

@@
 \section{Introduction and Motivation}
-Over the last decade investors have become increasingly concerned with extra factors outside of financials to address the performance of stocks.
+Asset pricing has a long history of models and theories governing how investor view the prices of assets.
+Over the last decade investors have become increasingly interested in  factors outside of traditional financial metrics when assessing the possability of an investments returns.Much of this enthusiasm has been based on ESG, Enviomental Social Goverance, investing that considers how firms respond to these factors as potential impetuses for excess performance.
 Research from Mckendsey has shown that there is a direct correlation between diversity inside of executive team leadership and company performance.
@@
 \section{Literature Review}
+to account for the ESG factor of a five factor famma french model, the authers used the S\&P ESG index as the high ESG index and the S\&P petroleum index as the low index creating a facotr by subtracting the retunrs from the high ESG portfolio from the low ESG portfolio.
+\cite{Gregory2020TheGP}
 \label{Literature}
@@
 The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices.
-Data on the ESG portfolio was extracted from compustat as the daily average returns for MSCI World ESG Focus Index.
+Data on the ESG portfolio and well as the petroleum portfolio was extracted from compustat as the daily average returns for MSCI World ESG Focus Index and S/&P petroleum index.
 Finally, daily Fama French factors for SMB, CMA,HML, and RMW were sourced from the Kenneth R.
 French - Data Library.
-The time period for the data for the paper is between June 2011 and December 2020.
-The segment of stock data is determined by the length of available performance for the MSCI World ESG Focus Index as this index was initiated in June 2011.
-Daily values were summed and collapsed to created monthly values for factor and individual security performance.
+An ESG factor, demoted as EMP (ESG minus petroleum), was created by taking the  MSCI World ESG Focus Index returns and subtracting them from the  S\&P petroleum index.
+Petroleum companies act as a stand in for the opposite of ESG as the  petroleum space is particularly inept to adapt to ESG  changes, a fact well documented in academic literature \cite{Frynas2005TheFD}.
+In addition, in section \ref{Robustness} for robustness I will repeat the analysis replacing the  S\&P petroleum index with S\&P Metals & Mining Select Industry Index as the opposed portfolio for the EMP factor.
+It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry \cite{Kapelus2002MiningCS}.
 
-we looking 830,891 observation for  12,194 unique stocks which gives an average of 68months of data
+ Data was collected between June 2011 and March 2021.
+The started was determined by the initiation of the MSCI ESG index.
+With the selected window, CRSP data for daily returns was collected on each available company.
+Daily returns for the individual companies as well as each of the Fama factors and the EMP factor was collapsed by month by suming the daily changes in each given month and year.
+The collapses resulted in 830,891 observation for 12,194 unique stocks which gives an average of 68months of data for each unique security.
+outlined in table \hl{1} is some of the summary statistics for the raw data.
 
@@
 \label{Methodology}
-As dicussed in the literature review there are numerous different models for assessing asset pricing and abnormal returns.
+As discussed in the literature review there are numerous different models for assessing asset pricing and abnormal returns.
 In this paper I will be using an extension on on Fama French's 2015 paper ``A five factor asset pricing model".
+This method of analysis is builds off of the famma french three factor model and relies on the 2 pass sort to assign a series of beta decile and size decile based portfolios.
+Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors do asses the abnormal returns.
+
+to determine the beta deciles for individual secutires a regression
+
+\begin{equation}
+    R_{it}^e = a_{i,t} + \beta_{i1,t}(R_{m,t}-R_{rf,t})+\beta_{i2,t}(R_{m,t-1}-R_{rf,t-1})+\epsilon_{i,t}
+\end{equation}
+\begin{equation}
+    \beta_{sum} = \sum_{k=1}^2\beta_{ik,t}
+\end{equation}
+
+Decile-sorted indexes based on
+market capitalization
+TO find the beta deciles
 
@@
 \begin{split}
-    R_{it}^e = a_i+b_i(R_{Mt}-R_{fr})+s_iSMB_t+ & h_iHML_t+ & r_iRMW_t+c_iCMA+e_i(R_{ESG t}-R_{rf})
+    R_{it}^e = a_i+b_i(R_{Mt}-R_{fr})+s_iSMB_t+ & h_iHML_t+ & r_iRMW_t+c_iCMA_t+e_iEMP_t
     \end{split}
@@
 \end{equation}
-In equation \eqref{six_factor_basic_reg} $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms, $e^e_{it}$ is the return of the individual secutiy minus the risk freee rate ($R_{it}-R_{rf}$)
+In equation \eqref{six_factor_basic_reg} $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms,$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index.
+finally, $R^e_{it}$ is the return of the individual secutiy minus the risk freee rate ($R_{it}-R_{rf}$)
 
@@
 \section{Robustness Checks}
+ %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
 \label{Robustness}
@@
 \label{Conclusion}
+
+\bibliography{Matt_Bib.bib}{}
+\bibliographystyle{plain}
+%\nocite{*}
+%\printbibliography
+
 \section{Code}

### 2021-04-11 Paper/Main.tex
starting evening grind

@@
 
-to determine the beta deciles for individual secutires a regression
+To determine the beta deciles for individual securities first a beta must be calculate for each security.
+To calculate the betas, a  simple regression outlined in equation \eqref{prerank_reg} was preformed.
+In this regression, a vector of  excess market returns in the current period and a vector of market returns from the previous period is regressed on a vector of current periods individual stock excess returns (i.e $R_{i,t}-R_{rf,t}$).
+The time movements for the data moved in month increments and the length of the vector ranged for 12 month to 60 months depending on the availability of data.
+Next, the beta values are summed together to get the pre-ranking beta value.
 
 \begin{equation}
-    R_{it}^e = a_{i,t} + \beta_{i1,t}(R_{m,t}-R_{rf,t})+\beta_{i2,t}(R_{m,t-1}-R_{rf,t-1})+\epsilon_{i,t}
+    R_{i,t}^e = a_{i,t} + \beta_{i1,t}(R_{m,t}-R_{rf,t})+\beta_{i2,t}(R_{m,t-1}-R_{rf,t-1})+\epsilon_{i,t}
+    \label{prerank_reg}
 \end{equation}
@@
 \end{equation}
+
+Before assigning the deciles a data cleaning process was applied.
+A small portion of prerankning beta's are far larger than would be expected, as such these vales were removed from the sample by removing top and bottom 4\% of companies by preranking beta.
+A similar data cleaning process was also conducted in the Fama french five factor model.
+In addition to removing the very large and very small preranking beta values I also removed the companies with market capitalization less than 10M dollars in market capitalization.
+After sorting by the beta decile's size deciles were asigned based on the CRSP NYSE/NYSE American/NASDAQ market capitalization.
 

### 2021-04-12 Code/cross_sectional_matlab.m
lll

@@ -9,7 +9,8 @@ smb = A{:,6};
 rmw = A{:,7};
 cma = A{:,8};
 esg_factor = A{:,9};
-time_index = A{:,10};
+market_factor = A{:,10};
+time_index = A{:,11};
 
 %% the show!!
 out_for_reg_1 = [];
@@ -19,15 +20,18 @@ for i = 1:length(porflio_number)-1
     current_stock = porflio_number(i);
     next_stock =porflio_number(i+1);
     
-    if current_stock == 100
+    if current_stock == 10
         returns_i = expected_return_stock(i-counter:end); 
         hml_i = hml(i-counter:end);
         smb_i = smb(i-counter:end);
         rmw_i= rmw(i-counter:end);
         cma_i = cma(i-counter:end);
         esg_factor_i = esg_factor(i-counter:end);
+        market_factor_i =market_factor(i-counter:end);
         factors = [hml_i smb_i rmw_i cma_i  esg_factor_i];
-        reg_1 = regress(returns_i, factors);
+        factors_w_market = [market_factor_i hml_i smb_i rmw_i cma_i  esg_factor_i];
+        %reg_1 = regress(returns_i, factors);
+        reg_1 = regress(returns_i, factors_w_market);
         out_for_reg_1 = [out_for_reg_1 reg_1];
         break
     end
@@ -37,14 +41,18 @@ for i = 1:length(porflio_number)-1
         counter = counter+1;
     else
         if counter >2
-        returns_i = expected_return_stock(i-counter:i); 
+        returns_i = expected_return_stock(i-counter:i);
+         
         hml_i = hml(i-counter:i);
         smb_i = smb(i-counter:i);
         rmw_i= rmw(i-counter:i);
         cma_i = cma(i-counter:i);
         esg_factor_i = esg_factor(i-counter:i);
+        market_factor_i =market_factor(i-counter:i);
         factors = [hml_i smb_i rmw_i cma_i esg_factor_i];
-        reg_1 = regress(returns_i, factors);
+        factors_w_market = [market_factor_i hml_i smb_i rmw_i cma_i  esg_factor_i];
+        %reg_1 = regress(returns_i, factors);
+        reg_1 = regress(returns_i, factors_w_market);
         out_for_reg_1 = [out_for_reg_1 reg_1];
         counter = 0;
         end 
@@ -58,7 +66,8 @@ out_for_reg_2 =[];
[…]

### 2021-04-12 Code/stage_1_pre_rank.m

@@ -24,10 +24,10 @@ mmonth = A{:,11};
 
 upper_counter =0;
 lower_counter = 0;
-pre_rank_matrix = nan(length(stock_symbols_refrences),13); %if I wanted individual beta we would extend this to 7
-for i =1:length(stock_symbols_refrences)-1
+pre_rank_matrix = nan(length(stock_symbols_refrences),13); %adjust the last term if you want to carry more values
+for i =2:length(stock_symbols_refrences);
     current_stock = stock_symbols_refrences(i);
-    next_stock =stock_symbols_refrences(i+1);
+    next_stock =stock_symbols_refrences(i-1);
     r_and_m_name = [current_stock individual_stock_return(i) ESG_returns(i) market_reutrns(i) cma(i) hml(i) smb(i) rmw(i) individual_stock_market_cap(i) cal_year(i) mmonth(i)];%year and month and name
     
     if current_stock == next_stock
@@ -36,9 +36,9 @@ for i =1:length(stock_symbols_refrences)-1
                 upper_counter = n_upper;
             end
             %indexing seems sus!! 4/9/21
-            market = [market_reutrns(1+i-n_lower-upper_counter:i+1) market_reutrns(i-n_lower-upper_counter:i)];
-            ESG_retun_data = [ESG_returns(1+i-n_lower-upper_counter:i+1) ESG_returns(i-n_lower-upper_counter:i)];
-            individual = [individual_stock_return(1+i-n_lower-upper_counter:i+1)]; 
+            market = [market_reutrns(i-n_lower-upper_counter:i) market_reutrns(i-n_lower-upper_counter-1:i-1)];
+            ESG_retun_data = [ESG_returns(i-n_lower-upper_counter:i) ESG_returns(i-n_lower-upper_counter-1:i-1)];
+            individual = [individual_stock_return(i-n_lower-upper_counter:i)]; 
             beta_market = regress(individual , market)';
             beta_ESG = regress(individual , ESG_retun_data)';
             sum_o_beta_market = sum(beta_market);

### 2021-04-13 Code/cross_sectional_matlab.m
lll

@@ -1,7 +1,7 @@
 %Matt Chistolini
-%Last edited 4/10/21
+%Last edited 4/12/21
 clear; clc;
-A = readtable("data/out_of_stata_six_factor.csv");
+A = readtable("data/size_porflios_and_returns.csv");
 porflio_number = A{:,3};
 expected_return_stock = A{:,4};
 hml = A{:,5};
@@ -28,10 +28,10 @@ for i = 1:length(porflio_number)-1
         cma_i = cma(i-counter:end);
         esg_factor_i = esg_factor(i-counter:end);
         market_factor_i =market_factor(i-counter:end);
-        factors = [hml_i smb_i rmw_i cma_i  esg_factor_i];
+        %factors = [hml_i smb_i rmw_i cma_i  esg_factor_i];
         factors_w_market = [market_factor_i hml_i smb_i rmw_i cma_i  esg_factor_i];
         %reg_1 = regress(returns_i, factors);
-        reg_1 = regress(returns_i, factors_w_market);
+        reg_1 = fitlm(returns_i, factors_w_market)
         out_for_reg_1 = [out_for_reg_1 reg_1];
         break
     end
@@ -49,10 +49,12 @@ for i = 1:length(porflio_number)-1
         cma_i = cma(i-counter:i);
         esg_factor_i = esg_factor(i-counter:i);
         market_factor_i =market_factor(i-counter:i);
-        factors = [hml_i smb_i rmw_i cma_i esg_factor_i];
+        %factors = [hml_i smb_i rmw_i cma_i esg_factor_i];
         factors_w_market = [market_factor_i hml_i smb_i rmw_i cma_i  esg_factor_i];
+        tabl = table(market_factor_i, hml_i, smb_i, rmw_i, cma_i,  esg_factor_i,returns_i)
         %reg_1 = regress(returns_i, factors);
-        reg_1 = regress(returns_i, factors_w_market);
+        %reg_1 = regress(returns_i, factors_w_market);
+        %reg_1 = regress(returns_i, factors_w_market);
         out_for_reg_1 = [out_for_reg_1 reg_1];
         counter = 0;
         end 

### 2021-04-13 Code/prelim_beta_weigh_regs.m

@@ -1,0 +1,20 @@
+%Matt Chistolini
+%Last edited 4/12/21
+clear; clc;
+%importing data
+A = readtable("data/size_porflios_and_returns.csv");
+%
+
+out_for_reg_1 = stage_one_reg(A,1)
+out_for_reg_total_coff = stage_one_reg(A,2)
+out_for_reg_SE = stage_one_reg(A,3)
+%%
+f = stage_two_reg(A,out_for_reg_1);
+%% Mean STD
+
+time_index = A{:,11};
+time_length = max(time_index);
+
+beta_avg = mean(f')*100 % in percent 
+beta_std = std(f')./sqrt(time_length)
+

### 2021-04-13 Code/stage_one_reg.m

@@ -1,0 +1,82 @@
+function f = stage_one_reg(A,output_option)
+%upgrade to MLS https://www.mathworks.com/help/stats/fitlm.html
+%defining parameters
+porflio_number = A{:,3};
+expected_return_stock = A{:,4};
+hml = A{:,5};
+smb = A{:,6};
+rmw = A{:,7};
+cma = A{:,8};
+esg_factor = A{:,9};
+market_factor = A{:,10};
+time_index = A{:,11};
+max_port = max(porflio_number);
+
+out_for_reg_1 = [];
+all_coff = [];
+all_SE= [];
+counter = 0;
+for i = 1:length(porflio_number)-1
+    
+    current_stock = porflio_number(i);
+    next_stock =porflio_number(i+1);
+    
+    if current_stock == max_port
+        returns_i = expected_return_stock(i-counter:end); 
+        hml_i = hml(i-counter:end);
+        smb_i = smb(i-counter:end);
+        rmw_i= rmw(i-counter:end);
+        cma_i = cma(i-counter:end);
+        esg_factor_i = esg_factor(i-counter:end);
+        market_factor_i =market_factor(i-counter:end);
+        
+        %making reg
+        factors_w_market_tbl = table(market_factor_i, hml_i, smb_i, rmw_i, cma_i,  esg_factor_i, returns_i);
+        reg_total_disp = fitlm(factors_w_market_tbl);
+        cofficents_include_a= table2array(reg_total_disp.Coefficients(:,1));
+        cofficents_no_a = cofficents_include_a(2:end);
+        SE_individual = table2array(reg_total_disp.Coefficients(:,2));
+        out_for_reg_1 = [out_for_reg_1 cofficents_no_a];
+        all_SE = [all_SE  SE_individual];
+        all_coff =[all_coff cofficents_include_a];
+        break
+    end
+    
+    if current_stock == next_stock
+        
+        counter = counter+1;
+    else
+        if counter >2
+        returns_i = expected_return_stock(i-counter:i);
+         
+        hml_i = hml(i-counter:i);
+        smb_i = smb(i-counter:i);
+        rmw_i= rmw(i-counter:i);
+        cma_i = cma(i-counter:i);
+        esg_factor_i = esg_factor(i-counter:i);
+        market_factor_i =market_factor(i-counter:i);
+        
+        %running reg
[…]

### 2021-04-13 Code/stage_two_reg.m

@@ -1,0 +1,15 @@
+function f = stage_two_reg(A,out_for_reg_1)
+time_index = A{:,11};
+expected_return_stock = A{:,4};
+time_length = max(time_index);
+out_for_reg_2 =[];
+counter = 0;
+for t = 1:1:time_length
+    counter=counter+1;
+    index_values = find(time_index ==t);
+    
+    reg_2 = regress(expected_return_stock(index_values),out_for_reg_1'); 
+    out_for_reg_2 = [out_for_reg_2 reg_2];
+end   
+f = out_for_reg_2;
+end

### 2021-04-14 Code/prelim_beta_weigh_regs.m
lll 2

@@ -9,7 +9,8 @@ out_for_reg_1 = stage_one_reg(A,1)
 out_for_reg_total_coff = stage_one_reg(A,2)
 out_for_reg_SE = stage_one_reg(A,3)
 %%
-f = stage_two_reg(A,out_for_reg_1);
+
+f = stage_two_reg(A,out_for_reg_1,"SE")
 %% Mean STD
 
 time_index = A{:,11};

### 2021-04-14 Code/stage_two_reg.m

@@ -1,15 +1,31 @@
-function f = stage_two_reg(A,out_for_reg_1)
+function f = stage_two_reg(A,out_for_reg_1,options)
 time_index = A{:,11};
 expected_return_stock = A{:,4};
 time_length = max(time_index);
 out_for_reg_2 =[];
+out_for_reg_2_w_a=[];
+all_SE=[];
 counter = 0;
 for t = 1:1:time_length
     counter=counter+1;
     index_values = find(time_index ==t);
-    
-    reg_2 = regress(expected_return_stock(index_values),out_for_reg_1'); 
-    out_for_reg_2 = [out_for_reg_2 reg_2];
+    LLL=out_for_reg_1';
+    tbl = table (LLL(:,1),LLL(:,2),LLL(:,3),LLL(:,4),LLL(:,5), LLL(:,6), expected_return_stock(index_values));
+    reg_2 = fitlm(tbl);
+    cofficents_include_a= table2array(reg_2.Coefficients(:,1));
+    SE_individual = table2array(reg_2.Coefficients(:,2));
+    cofficents_no_a = cofficents_include_a(2:end);
+    out_for_reg_2 = [out_for_reg_2 cofficents_no_a];
+    out_for_reg_2_w_a= [out_for_reg_2_w_a cofficents_include_a];
+    all_SE = [all_SE  SE_individual];
 end   
-f = out_for_reg_2;
+if options == "a_plz"
+    f = out_for_reg_2_w_a;
+elseif options == "no_a"
+    f = out_for_reg_2;
+elseif  options == "SE"
+    f =all_SE;
 end
+
+end
+

### 2021-04-14 Paper/Main.tex

@@
 \section{Introduction and Motivation}
-Asset pricing has a long history of models and theories governing how investor view the prices of assets.
-Over the last decade investors have become increasingly interested in  factors outside of traditional financial metrics when assessing the possability of an investments returns.Much of this enthusiasm has been based on ESG, Enviomental Social Goverance, investing that considers how firms respond to these factors as potential impetuses for excess performance.
-Research from Mckendsey has shown that there is a direct correlation between diversity inside of executive team leadership and company performance.
-A rising threat of global climate related issues and workplace culture to determine top talent have all become extremely important for company management to consider on the basis of long term performance.
+Asset pricing has a long history of models and theories governing how investors make decisions.
+Over the last decade, investors have become increasingly interested in factors outside of traditional financial metrics when assessing the possibility of an investment's returns.
+Much of this enthusiasm has been based on an ESG, Environmental Social Governance, framework.
+ESG investing considers how firms respond to these factors as potential impetuses for excess performance.
+This increased attention is especially visible in Morgan Stanley's 2019 survey of the ESG landscape.
+Morgan Standley  found that 75\% of assessment managers say that "their firm has adopted sustainable investing." Only two years prior, in 2016, just 10\% of surveyed asset managed had adopted sustainable investing, demonstrating the rapid speed of adoption for ESG investing analysis \cite{morgan_stanley_sustainable_2019}.
+Practitioners have utilized the ESG framework to reduce risk in their investments while also finding abnormal returns.
 
-this paper will be broken up into five  sections.
-\textit{Section \ref{Literature}} will provide a  brief literature review of papers looking abnormal performance around ESG criteria as well as discussion of possible asset pricing models to capture ESG factors in performance.
-Next in  \textit{ Section \ref{Data}}  I will discuss the data used in this papers as well as the data cleaning process and summary statistics.
+Global climate change presents a host of possible and sizable financial risks from increased water levels, which could threaten major global cities' safety, including Miami, Honk Konk, and Shanghai, with the potential to displace 100's of millions of people \cite{holder_three-degree_nodate}.
+Global climate change is also shifting the landscape for global agricultural production as more extreme weather conditions, including droughts, pose problems to the most fundamental building block of economic activity.
+In addition to risk mitigation, investors look to utilize ESG principle for superior returns.
+A research report published by McKinsey and Company linked gender diversity on executive teams to higher profitability and superior value creation \cite{mckinsey_diversity}.
+
+In addition to risk investing professionals, ESG investing is appealing to a younger set of investors interested in a plethora of global issues from global climate change to diversity and inclusion in the workplace.
+A study conducted by Morgan Stanly found that millennials investors were twice as likely as the general investing population to invest in a company with social or environmental goals.
+
+This new wave of investor sentiment has created strong tailwinds for sustainable investment projects.
+In 2020 alone, fund flows into sustainable funds grew by 51 billion dollars which is up over 100\% for the 2019 fund flow into sustainable investment products of about 21, placing total assets under management in sustainable products at over 250B.
+Fund flows into sustainable products for 2020 accounted for a little less a quarter of all US mutual fund flow and about one-fifteenth of all ETF fund flow according to morningstar \cite{monring_star_ESG}.
+
+To address the impact of ESG factors on stock preformance, this paper is broken into five sections.
+\textit{Section \ref{Literature}} will provide a  brief literature review of papers looking at abnormal performance around ESG criteria as well as discussion of asset pricing models to capture ESG factors in performance.
+Next \textit{ Section \ref{Data}}  will provide a brief discussion of the data used in this paper as well as the data cleaning process and summary statistics.
 Following the Data, \textit{Section \ref{Methodology}} will breakdown of the methodology for the paper.
@@
 \section{Literature Review}
+This section will be broke into two components, we will first look at theoretical models for asset purchasing and then we will introduce some of the mythology for introducing a EGS factor into the asset pricing framework.
 to account for the ESG factor of a five factor famma french model, the authers used the S\&P ESG index as the high ESG index and the S\&P petroleum index as the low index creating a facotr by subtracting the retunrs from the high ESG portfolio from the low ESG portfolio.
@@
 
+*dicuss how we collapse the prolfio from 10 to 2 size
+
 Decile-sorted indexes based on

### 2021-04-14 Paper/Main.tex
Updates from Overleaf

@@
 Daily returns for the individual companies as well as each of the Fama factors and the EMP factor was collapsed by month by suming the daily changes in each given month and year.
-The collapses resulted in 830,891 observation for 12,194 unique stocks which gives an average of 68months of data for each unique security.
-outlined in table \hl{1} is some of the summary statistics for the raw data.
+The collapses resulted in 830,889 observation for 12,194 unique stocks which gives an average of 68 months of data for each unique security.
+Outlined in table are  some of the summary statistics for the raw data.
 
-to reduce some of the tail behavior we dropped the top and bottom  4\% of pre ranked beta values as well as companies with market values less tham 10 million
+ \begin{center}
+     \textbf{Table 1}\\
+ \input{Code/data/desc1}
+ \end{center}
+
 \section{Methodology}

### 2021-04-15 Paper/Main.tex
just pushing b/c we have done alot

@@
 \label{Data}
-The data used for this paper was soured three databases.
+The data used for this paper was sourced from three datasets.
 The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices.
-Data on the ESG portfolio and well as the petroleum portfolio was extracted from compustat as the daily average returns for MSCI World ESG Focus Index and S/&P petroleum index.
-Finally, daily Fama French factors for SMB, CMA,HML, and RMW were sourced from the Kenneth R.
+Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily average returns for the MSCI World ESG Focus Index and S/&P petroleum index, respectively.
+Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R.
 French - Data Library.
-An ESG factor, demoted as EMP (ESG minus petroleum), was created by taking the  MSCI World ESG Focus Index returns and subtracting them from the  S\&P petroleum index.
-Petroleum companies act as a stand in for the opposite of ESG as the  petroleum space is particularly inept to adapt to ESG  changes, a fact well documented in academic literature \cite{Frynas2005TheFD}.
-In addition, in section \ref{Robustness} for robustness I will repeat the analysis replacing the  S\&P petroleum index with S\&P Metals & Mining Select Industry Index as the opposed portfolio for the EMP factor.
+An ESG factor, denoted as EMP (ESG minus petroleum), was created by taking the  MSCI World ESG Focus Index returns and subtracting them from the  S\&P petroleum index.
+Petroleum companies act as a stand-in for the opposite of ESG as the petroleum space is particularly inept at adapting to ESG  changes, a fact well documented in academic literature \cite{Frynas2005TheFD}.
+In addition, in section \ref{Robustness} for robustness I will repeat the analysis replacing the  S\&P petroleum index with S\&P Metals & Mining Select Industry Index as the contra ESG portfolio for the EMP factor.
 It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry \cite{Kapelus2002MiningCS}.
 
- Data was collected between June 2011 and March 2021.
+ Data points were collected between June 2011 and March 2021.
 The started was determined by the initiation of the MSCI ESG index.
-With the selected window, CRSP data for daily returns was collected on each available company.
-Daily returns for the individual companies as well as each of the Fama factors and the EMP factor was collapsed by month by suming the daily changes in each given month and year.
-The collapses resulted in 830,889 observation for 12,194 unique stocks which gives an average of 68 months of data for each unique security.
-Outlined in table are  some of the summary statistics for the raw data.
+With the selected window, CRSP data for daily returns were collected on each available company.
+Daily returns for the individual companies as well as each of the Fama factors and the EMP factor was collapsed by month by summing the daily changes in each given month and year.
+The collapses resulted in 830,891 observations for 12,194 unique stocks, which gives an average of 68months of data for each unique security—outlined in table 1are some of the summary statistics for the raw data.
 
@@
 In this paper I will be using an extension on on Fama French's 2015 paper ``A five factor asset pricing model".
-This method of analysis is builds off of the famma french three factor model and relies on the 2 pass sort to assign a series of beta decile and size decile based portfolios.
-Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors do asses the abnormal returns.
+This method of analysis is builds off of the famma french three factor model and relies on the 2 pass sort with an assign series of beta decile and size decile based portfolios.
+Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors to assess relational interdependency.
 
-To determine the beta deciles for individual securities first a beta must be calculate for each security.
-To calculate the betas, a  simple regression outlined in equation \eqref{prerank_reg} was preformed.
-In this regression, a vector of  excess market returns in the current period and a vector of market returns from the previous period is regressed on a vector of current periods individual stock excess returns (i.e $R_{i,t}-R_{rf,t}$).
-The time movements for the data moved in month increments and the length of the vector ranged for 12 month to 60 months depending on the availability of data.
-Next, the beta values are summed together to get the pre-ranking beta value.
+In the preliminary analysis of the ESG factor I sort securties  on beta as well as size deciles.
+Each security will be assigned a decile for beta and size according the the CRSP /NYSE american beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size.
+After sorting each of the securities by Beta and size decile  300,264 of the original 830,889 data points had matches for both portfolios.
+To capture the effect of size and each of the Fama French factors and ESG factor I ran the reguression in equation \eqref{six_factor_basic_reg} on the cleaned data set.
 
 \begin{equation}
-    R_{i,t}^e = a_{i,t} + \beta_{i1,t}(R_{m,t}-R_{rf,t})+\beta_{i2,t}(R_{m,t-1}-R_{rf,t-1})+\epsilon_{i,t}
-    \label{prerank_reg}
+\begin{split}
+    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ & h_iHML_t+ & r_iRMW_t+c_iCMA_t+e_iEMP_t
+    \end{split}
+    \label{six_factor_basic_reg}
 \end{equation}
+Where:
+\begin{center}
+    $R_{i,t}^e = R_{it}-R_{rf,t}$
+\end{center}
+
+***Dicuss the second pass***
+
 \begin{equation}
-    \beta_{sum} = \sum_{k=1}^2\beta_{ik,t}
+    E[R^e_{t}]= b_t\lambda_1+s_t\lambda_2 +h_t \lambda_3 +r_t\lambda_4 +c_t \lambda_4 + e_t \lambda_5
 \end{equation}
+In equation \eqref{six_factor_basic_reg}
 
-Before assigning the deciles a data cleaning process was applied.
-A small portion of prerankning beta's are far larger than would be expected, as such these vales were removed from the sample by removing top and bottom 4\% of companies by preranking beta.
-A similar data cleaning process was also conducted in the Fama french five factor model.
-In addition to removing the very large and very small preranking beta values I also removed the companies with market capitalization less than 10M dollars in market capitalization.
-After sorting by the beta decile's size deciles were asigned based on the CRSP NYSE/NYSE American/NASDAQ market capitalization.
+\section{Results}
+\label{Results}
+table 3 in represents a matrix of values for the first pass in the fama french two pass sort, described in equation \eqref{six_factor_basic_reg}.
+The column axis is the distinct size deciles where 1 is the smallest market cap companies and 10 is the declie with the largest market cap companies.
+The row axis includes estimated coefficients for the intercept, five Fama French factors and the EMS factor.
+$HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms,$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index.
+finally, the intercept is the excess returns for the given decile.The results in this table demonstrate that for HML, CMA and EMP factor there is a size based relationship where in the larger the size decile the smaller the effect is on the factor.
+In addition, the EMP factor also exhibits a  a consistent negative relationship between the EGS factor and returns.
+This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra EGS counterparts.
 
-*dicuss how we collapse the prolfio from 10 to 2 size
+Moving onto the 
+Analysing our results from the methology  stages of the two pass sort we would expect to see
 
-Decile-sorted indexes based on
-market capitalization
-TO find the beta deciles
+\begin{center}
+\input{Paper/prelim_size}\\
+\textbf{Table 3:} Size decile by estimated coefficients for the  five fama french factors and EMP factor
+\end{center}
+\begin{center}
+\input{Paper/p_vals_size_prelim}\\
+\textbf{Table 4:} Size decile by p-values for table 3
+
+\end{center}
+
+\begin{center}
+\input{Paper/prelim_beta}\\
+\textbf{Table 5:} beta decile by estimated coefficients for the  five fama french factors and EMP factor
+\end{center}
 
@@
 For this purpose we will use the MSCI ESG all world index.
-
-\begin{equation}
-\begin{split}
-    R_{it}^e = a_i+b_i(R_{Mt}-R_{fr})+s_iSMB_t+ & h_iHML_t+ & r_iRMW_t+c_iCMA_t+e_iEMP_t
-    \end{split}
-    \label{six_factor_basic_reg}
-\end{equation}
-In equation \eqref{six_factor_basic_reg} $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms,$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index.
-finally, $R^e_{it}$ is the return of the individual secutiy minus the risk freee rate ($R_{it}-R_{rf}$)
-
-\section{Results}
-\label{Results}
-
 \section{Robustness Checks}
@@
 \bibliographystyle{plain}
+
+\newgeometry{margin=.15in}
+\input{Code/data/size_summary_stat_size}\\
+
+\vspace{1cm}
+\textbf{Table 2:}things\\
+\vspace{1cm}
+
+\clearpage
+\restoregeometry
+
 %\nocite{*}
@@
 
-\section{Code}
-\input{Paper/Code_files}
+%\section{Code}
+%\input{Paper/Code_files}
 \end{document}

### 2021-04-15 Paper/junk_maybe.tex
just pushing b/c we have done alot

To determine the beta deciles for individual securities first a beta must be calculate for each security. To calculate the betas, a  simple regression outlined in equation \eqref{prerank_reg} was preformed. In this regression, a vector of  excess market returns in the current period and a vector of market returns from the previous period is regressed on a vector of current periods individual stock excess returns (i.e $R_{i,t}-R_{rf,t}$). The time movements for the data moved in month increments and the length of the vector ranged for 12 month to 60 months depending on the availability of data. Next, the beta values are summed together to get the pre-ranking beta value.

\begin{equation}
    R_{i,t}^e = a_{i,t} + \beta_{i1,t}(R_{m,t}-R_{rf,t})+\beta_{i2,t}(R_{m,t-1}-R_{rf,t-1})+\epsilon_{i,t}
    \label{prerank_reg}
\end{equation}
\begin{equation}
    \beta_{sum} = \sum_{k=1}^2\beta_{ik,t}
\end{equation}

### 2021-04-15 Paper/Main.tex
mad edits

@@
 After sorting each of the securities by Beta and size decile  300,264 of the original 830,889 data points had matches for both portfolios.
-To capture the effect of size and each of the Fama French factors and ESG factor I ran the reguression in equation \eqref{six_factor_basic_reg} on the cleaned data set.
+To estimate the factor sensitivities base on the discribed decile groups I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.
 
@@
 \end{equation}
-Where:
+
 \begin{center}
+Where:\\
     $R_{i,t}^e = R_{it}-R_{rf,t}$
@@
 
-***Dicuss the second pass***
+Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two pass sort.
+The second pass of the two pass sort is  described in the equation  \eqref{second_pass} and is used to obtain the risk premia.
+To consolidate the the risk premium across time a simple average as described in equation \eqref{summing_lamda} is used to collapse the risk premia for each of the given factors.
 
@@
     E[R^e_{t}]= b_t\lambda_1+s_t\lambda_2 +h_t \lambda_3 +r_t\lambda_4 +c_t \lambda_4 + e_t \lambda_5
+    \label{second_pass}
 \end{equation}
-In equation \eqref{six_factor_basic_reg}
+
+\begin{equation}
+    \frac{1}{T}\sum^T_{t=1}\lambda_{i,t}
+    \label{summing_lamda}
+\end{equation}
+\begin{center}
+    where:\\
+    T: is equal to the number of months in sample
+\end{center}
 
@@
 \label{Results}
-table 3 in represents a matrix of values for the first pass in the fama french two pass sort, described in equation \eqref{six_factor_basic_reg}.
+Table 3 in represents a matrix of values for the first pass in the fama french two pass sort, described in equation \eqref{six_factor_basic_reg}.
 The column axis is the distinct size deciles where 1 is the smallest market cap companies and 10 is the declie with the largest market cap companies.
@@
 In addition, the EMP factor also exhibits a  a consistent negative relationship between the EGS factor and returns.
-This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra EGS counterparts.
+This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts.
+Moving onto the second step in the two pass sort to asess the risk premia,
 
-Moving onto the 
-Analysing our results from the methology  stages of the two pass sort we would expect to see
-
+Moving onto the analysing the results of the second stage in the two pass sort we would expect to see that   
+% size cofficents
 \begin{center}
@@
 \end{center}
+% size p vals
 \begin{center}
@@
 \textbf{Table 4:} Size decile by p-values for table 3
-
 \end{center}
-
+% beta cofficents
 \begin{center}
@@
 \textbf{Table 5:} beta decile by estimated coefficients for the  five fama french factors and EMP factor
+\end{center}
+% beta p vals
+\begin{center}
+\input{Paper/p_vals_beta_prelim}\\
+\textbf{Table 6:} Size decile by p-values for table 5
 \end{center}
@@
 \vspace{1cm}
-\textbf{Table 2:}things\\
+\textbf{Table 2:} This table shows many things\\
 \vspace{1cm}

### 2021-04-15 Code/prelim_beta_weigh_regs.m
YOOO

@@ -3,19 +3,43 @@
 clear; clc;
 %importing data
 A = readtable("data/size_porflios_and_returns.csv");
-%
+%Step 1
 
-out_for_reg_1 = stage_one_reg(A,1)
-out_for_reg_total_coff = stage_one_reg(A,2)
-out_for_reg_SE = stage_one_reg(A,3)
-%%
+out_for_reg_1 = stage_one_reg(A,1);
 
+out_for_reg_total_coff = stage_one_reg(A,2);
+out_for_reg_SE = stage_one_reg(A,3);
+% SE is actually p-valus ATM (4/14/21)
 f = stage_two_reg(A,out_for_reg_1,"SE")
-%% Mean STD
-
+%Step 2
 time_index = A{:,11};
 time_length = max(time_index);
-
-beta_avg = mean(f')*100 % in percent 
+%averages and shit
+beta_avg = mean(f')' %not precent basedintercept
+market-rf
+hml
+smb
+rmw
+cma
+EMP
 beta_std = std(f')./sqrt(time_length)
+
+%% Same thing with beta
+A = readtable("data/beta_porflios_and_returns.csv");
+%Step 1
+
+out_for_reg_1 = stage_one_reg(A,1)
+
+out_for_reg_total_coff = stage_one_reg(A,2)
+out_for_reg_SE = stage_one_reg(A,3)
+
+f = stage_two_reg(A,out_for_reg_1,"SE")
+%Step 2
+time_index = A{:,11};
+time_length = max(time_index);
+%averages and shit
+beta_avg = mean(f')' % not in percent 
+beta_std = std(f')./sqrt(time_length)
+
+
 

### 2021-04-15 Code/stage_one_reg.m

@@ -35,9 +35,9 @@ for i = 1:length(porflio_number)-1
         reg_total_disp = fitlm(factors_w_market_tbl);
         cofficents_include_a= table2array(reg_total_disp.Coefficients(:,1));
         cofficents_no_a = cofficents_include_a(2:end);
-        SE_individual = table2array(reg_total_disp.Coefficients(:,2));
+        P_Vals_individual = table2array(reg_total_disp.Coefficients(:,4));
         out_for_reg_1 = [out_for_reg_1 cofficents_no_a];
-        all_SE = [all_SE  SE_individual];
+        all_SE = [all_SE  P_Vals_individual];
         all_coff =[all_coff cofficents_include_a];
         break
     end
@@ -61,9 +61,9 @@ for i = 1:length(porflio_number)-1
         reg_total_disp = fitlm(factors_w_market_tbl);
         cofficents_include_a= table2array(reg_total_disp.Coefficients(:,1));
         cofficents_no_a = cofficents_include_a(2:end);
-        SE_individual = table2array(reg_total_disp.Coefficients(:,2));
+        P_Vals_individual = table2array(reg_total_disp.Coefficients(:,2));
         out_for_reg_1 = [out_for_reg_1 cofficents_no_a];
-        all_SE = [all_SE  SE_individual];
+        all_SE = [all_SE  P_Vals_individual];
         all_coff =[all_coff cofficents_include_a];
         counter = 0;
         end 

### 2021-04-15 Paper/Main.tex
why not another push

@@
 This increased attention is especially visible in Morgan Stanley's 2019 survey of the ESG landscape.
-Morgan Standley  found that 75\% of assessment managers say that "their firm has adopted sustainable investing." Only two years prior, in 2016, just 10\% of surveyed asset managed had adopted sustainable investing, demonstrating the rapid speed of adoption for ESG investing analysis \cite{morgan_stanley_sustainable_2019}.
+Morgan Stanley  found that 75\% of assessment managers say that "their firm has adopted sustainable investing." Only two years prior, in 2016, just 10\% of surveyed asset managed had adopted sustainable investing, demonstrating the rapid speed of adoption for ESG investing analysis \cite{morgan_stanley_sustainable_2019}.
 Practitioners have utilized the ESG framework to reduce risk in their investments while also finding abnormal returns.
@@
 The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices.
-Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily average returns for the MSCI World ESG Focus Index and S/&P petroleum index, respectively.
+Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily returns for the MSCI World ESG Focus Index and S\&P petroleum index, respectively.
 Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R.
 French - Data Library.
-An ESG factor, denoted as EMP (ESG minus petroleum), was created by taking the  MSCI World ESG Focus Index returns and subtracting them from the  S\&P petroleum index.
-Petroleum companies act as a stand-in for the opposite of ESG as the petroleum space is particularly inept at adapting to ESG  changes, a fact well documented in academic literature \cite{Frynas2005TheFD}.
-In addition, in section \ref{Robustness} for robustness I will repeat the analysis replacing the  S\&P petroleum index with S\&P Metals & Mining Select Industry Index as the contra ESG portfolio for the EMP factor.
+
+An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index.
+Petroleum companies act as a stand-in for a contra ESG proflio in a similar way to the what Fama french do to isolate the specific risk component.
+The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis.
+In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}.
+Although the Petroleum industry has major environmental, risks there is less social orientated risks as the index \$VICE which is an index of alcohol, drug, gaming and betting industry might provide a better exposure to the broad based effects of social ESG risk factors as a contra ESG portfolio.
+However \$Vice has only traded for two years, as such the data is far less robust.
+Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the  In addition, in section \ref{Robustness} for robustness I will repeat the analysis replacing the  S\&P petroleum index with S\&P Metals \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor.
 It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry \cite{Kapelus2002MiningCS}.
+
+\begin{equation}
+    \label{making_EMP}
+    EMP = R_{\text{World ESG Focus Index}}- R_{\text{S\&P petroleum index}}
+\end{equation}
 
@@
 
-In the preliminary analysis of the ESG factor I sort securties  on beta as well as size deciles.
+For this analysis I sort each security by month into decile groups for beta as well as size (based on market cap).
 Each security will be assigned a decile for beta and size according the the CRSP /NYSE american beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size.
@@
 \begin{split}
-    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ & h_iHML_t+ & r_iRMW_t+c_iCMA_t+e_iEMP_t
+    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t
     \end{split}
@@
 $HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms,$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index.
-finally, the intercept is the excess returns for the given decile.The results in this table demonstrate that for HML, CMA and EMP factor there is a size based relationship where in the larger the size decile the smaller the effect is on the factor.
+Finally, the intercept is the excess returns for the given decile.The results in this table demonstrate that for HML, CMA and EMP factor there is a size based relationship where in the larger the size decile the smaller the effect is on the factor.
 In addition, the EMP factor also exhibits a  a consistent negative relationship between the EGS factor and returns.
 This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts.
-Moving onto the second step in the two pass sort to asess the risk premia,
+Table 4 complements table 3 with a breakdown of significance for each coefficient in table 3.
+The low p-values in the EMP factor indicate that the coefficients on all but the largest and smallest size deciles are significant at the 5\% level.
+It is important to note that all of the factors are less significant that largest and small polios indicating that the data might have point at far ends which do not behave like the other deciles.
+Moving to table 6 and the second step in the two pass sort we can see the risk premia for each of the factors.
+All of the Fama French factors in the size decile two pass sort have positive risk premia.
+The EMP factor large than all of the other factors with a coefficient of approximately  16.8\%
 
-Moving onto the analysing the results of the second stage in the two pass sort we would expect to see that   
 % size cofficents
@@
 % beta cofficents
+
+Moving on to the beta decile based two pass sort the reuslts are quite similar to the size decile results.
+Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile where 1 is the portfolio comprised of the companies with lowest bet and 10 is the portfolio of companies  with the greatest beta values.
+Similar to the size decile the EMP factor exhabits a negative relations with regards to its sensitivity.
+However, unlike the size decile analyis there does not appear to be a signficant relationship between the  beta and the sensitivity of the factor.
+Although the largest impact is on the lowest decile the least impact is experienced on the 6th decile porfolio and the largest deciel is close to the middle in terms of sensitivity.
+Similar to the size decile analysis the EMP coefficients are all very significant with every decile having significance at the 5\% level.
+When considering the second phase of the two pass sort the similarities to the size decile continue where in each of the Fama French five factors exhabit a positive risk premia and the EMP also exhabits a positive relationship.
+the EMP factor again has the high coefficient with approximately 15.8\%.
+
 \begin{center}
@@
 \textbf{Table 6:} Size decile by p-values for table 5
+\end{center}
+% pass 2
+\begin{center}
+\input{Paper/pass_2}\\
+\textbf{Table 7:} second pass results for the size and beta decile portfolios
 \end{center}

### 2021-04-15 Paper/Main.tex
commit EON

@@
 An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index.
-Petroleum companies act as a stand-in for a contra ESG proflio in a similar way to the what Fama french do to isolate the specific risk component.
+Petroleum companies act as a stand-in for a contra ESG proflio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors.
 The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis.
 In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}.
-Although the Petroleum industry has major environmental, risks there is less social orientated risks as the index \$VICE which is an index of alcohol, drug, gaming and betting industry might provide a better exposure to the broad based effects of social ESG risk factors as a contra ESG portfolio.
+Although the Petroleum industry has major environmental, risks there is less social orientated risks in petrolium.
+One solution to this provlem would be to use the index \$VICE which tracks alcohol, drug, gaming and betting industry  as the  contra ESG portfolio.
 However \$Vice has only traded for two years, as such the data is far less robust.
-Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the  In addition, in section \ref{Robustness} for robustness I will repeat the analysis replacing the  S\&P petroleum index with S\&P Metals \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor.
-It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry \cite{Kapelus2002MiningCS}.
+Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor in Section \ref{Robustness}.
+It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry as has suffered from many human right violations which would significantly effects its social risk  \cite{Kapelus2002MiningCS}.
 
@@
 
- Data points were collected between June 2011 and March 2021.
+For this analysis, daily returns data was collected between June 2011 and March 2021.
 The started was determined by the initiation of the MSCI ESG index.
-With the selected window, CRSP data for daily returns were collected on each available company.
-Daily returns for the individual companies as well as each of the Fama factors and the EMP factor was collapsed by month by summing the daily changes in each given month and year.
-The collapses resulted in 830,891 observations for 12,194 unique stocks, which gives an average of 68months of data for each unique security—outlined in table 1are some of the summary statistics for the raw data.
+To collect daily individual stock return information CRSP data parsed from a larger dataset and then collapsed on month by summing each individual days returns to creat monthly returns.
+Similarly Fama French factors were collected on a daily basis and collapsed by month summing the daily returns for montly figures.
+Finally the EMP factor was also  constructed summing the daily returns for each much.
+All three of the datasets were merged using the date and month as the matching component.
+The culmination of merging and collapsing resulted in  830,891 observations for monthly returns over  12,194 unique stocks.
+The merged and cleaned dataset had an average of 68 months of data of individual security data for each unique security.
+Below in table 1 are the summary statistics for the raw data.
 
@@
 \label{Methodology}
-As discussed in the literature review there are numerous different models for assessing asset pricing and abnormal returns.
-In this paper I will be using an extension on on Fama French's 2015 paper ``A five factor asset pricing model".
-This method of analysis is builds off of the famma french three factor model and relies on the 2 pass sort with an assign series of beta decile and size decile based portfolios.
+As discussed in the literature review, there are numerous models for assessing asset pricing and abnormal returns.
+In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model".
+This analysis method is built on the Fama French three factor model and relies on the two-pass sort with an assigned series of beta decile and size decile-based portfolios.
 Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors to assess relational interdependency.
 
-For this analysis I sort each security by month into decile groups for beta as well as size (based on market cap).
-Each security will be assigned a decile for beta and size according the the CRSP /NYSE american beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size.
-After sorting each of the securities by Beta and size decile  300,264 of the original 830,889 data points had matches for both portfolios.
-To estimate the factor sensitivities base on the discribed decile groups I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.
+For this analysis, I sort each security into decile groups for beta and size (based on market cap).
+As firms' size and volatility are not constant functions, each firm is assigned a decile for each month depending on the most recent decile split.
+Each security will be assigned a decile for beta and size according to the CRSP /NYSE American beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size.
+After sorting each of the securities by beta and size decile, 300,264 of the original 830,889 data points had matches for both beta and size portfolios.
+To estimate the factor sensitivities base on the described decile groups, I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.
 
@@
 
-Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two pass sort.
-The second pass of the two pass sort is  described in the equation  \eqref{second_pass} and is used to obtain the risk premia.
-To consolidate the the risk premium across time a simple average as described in equation \eqref{summing_lamda} is used to collapse the risk premia for each of the given factors.
+Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort.
+The second pass of the two-pass sort is described in the equation  \eqref{second_pass} and is used to obtain the risk premia.
+To consolidate the risk premium across time, a simple average as described in equation \eqref{summing_lamda} is used to collapse the risk premia for each of the given factors.
 
@@
 \label{Results}
-Table 3 in represents a matrix of values for the first pass in the fama french two pass sort, described in equation \eqref{six_factor_basic_reg}.
-The column axis is the distinct size deciles where 1 is the smallest market cap companies and 10 is the declie with the largest market cap companies.
-The row axis includes estimated coefficients for the intercept, five Fama French factors and the EMS factor.
-$HML_{t}$ is the difference between the returns on diversified portfolios of high and low book to market stocks, $SMB_{t}$ i the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks, $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability,$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms,$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index.
-Finally, the intercept is the excess returns for the given decile.The results in this table demonstrate that for HML, CMA and EMP factor there is a size based relationship where in the larger the size decile the smaller the effect is on the factor.
-In addition, the EMP factor also exhibits a  a consistent negative relationship between the EGS factor and returns.
+Table 3 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}.
+The column axis is the distinct size deciles where 1 is the smallest market cap companies, and 10 is the decline with the largest market cap companies.
+The row axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor.
+$HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.
+$SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks.
+$RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability.
+$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms.
+$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index.
+Finally, the intercept is the excess returns for the given decile.
+The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor.
+In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns.
 This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts.
@@
 The low p-values in the EMP factor indicate that the coefficients on all but the largest and smallest size deciles are significant at the 5\% level.
-It is important to note that all of the factors are less significant that largest and small polios indicating that the data might have point at far ends which do not behave like the other deciles.
-Moving to table 6 and the second step in the two pass sort we can see the risk premia for each of the factors.
-All of the Fama French factors in the size decile two pass sort have positive risk premia.
-The EMP factor large than all of the other factors with a coefficient of approximately  16.8\%
+It is important to note that all of the other Fama French factors are also less significant In the largest and small portfolios.
+This trend might indicate that the data might have pointed at far ends which do not behave like the other deciles.
+Table 6 demonstrates the risk premia for each of the factors using the second step in the two-pass sort.
+All of the Fama French factors in the size decile two-pass sort have positive risk premia.
+The EMP factor large than all of the other factors, with a coefficient of approximately  16.8\%
 
@@
 
-Moving on to the beta decile based two pass sort the reuslts are quite similar to the size decile results.
-Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile where 1 is the portfolio comprised of the companies with lowest bet and 10 is the portfolio of companies  with the greatest beta values.
-Similar to the size decile the EMP factor exhabits a negative relations with regards to its sensitivity.
-However, unlike the size decile analyis there does not appear to be a signficant relationship between the  beta and the sensitivity of the factor.
-Although the largest impact is on the lowest decile the least impact is experienced on the 6th decile porfolio and the largest deciel is close to the middle in terms of sensitivity.
-Similar to the size decile analysis the EMP coefficients are all very significant with every decile having significance at the 5\% level.
-When considering the second phase of the two pass sort the similarities to the size decile continue where in each of the Fama French five factors exhabit a positive risk premia and the EMP also exhabits a positive relationship.
-the EMP factor again has the high coefficient with approximately 15.8\%.
+Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile.
+In this table,  the beta decile 1 comprises the companies with the lowest beta.
+Similarly, beta decile 10 is the portfolio of companies with the largest beta values.
+Similar to the size decile, the EMP factor exhibits a negative relation with regard to its sensitivity.
+However, unlike the size decile analysis, there does not appear to be a significant relationship between the beta decile and the factor's sensitivity.
+Although the most significant impact is on the lowest decile, the least impact is experienced on the 6th decile portfolio, and the largest decile is close to the middle in terms of sensitivity.
+Similar to the size decile analysis, the EMP coefficients are all very significant, with every decile having significance at the 5\% level.
+When considering the second phase of the two-pass sort, the similarities to the size decile continue wherein each of the Fama French five factors exhibit positive risk premia, and the EMP also exhibits a positive relationship.
+The EMP factor again has a high coefficient with approximately 15.8\%.
 

### 2021-04-25 Paper/Main.tex

@@
 \section{Literature Review}
-This section will be broke into two components, we will first look at theoretical models for asset purchasing and then we will introduce some of the mythology for introducing a EGS factor into the asset pricing framework.
-to account for the ESG factor of a five factor famma french model, the authers used the S\&P ESG index as the high ESG index and the S\&P petroleum index as the low index creating a facotr by subtracting the retunrs from the high ESG portfolio from the low ESG portfolio.
-\cite{Gregory2020TheGP}
 \label{Literature}
+This section will be broke into two components.
+The first section will adress  theoretical models for asset purchasing and then we will introduce some of the mythology for  EGS factor in the asset pricing framework.\\
+
+ \hl{add a section on the metholoogy for the 2 pass sort use} \cite{Kleibergen2013UnexplainedFA} and \cite{Fama1973RiskRA} \\
+The modern foundation of asset pricing began with the capital asset pricing model based on Sharp's 1964's work which used a regression analysis to relate market returns to individual security returns \cite{Sharpe1964CAPITALAP}.
+The CAPM model was later iterated on by Fama french in 1993 with the 3 factor, which considered several other terms to predict stock returns.
+Fama and French added size and book to market terms to the CAPM model.
+They found that these three coefficients created a more explanatory model which demonstrated the relationship that value companies(low book to market) and small companies experience larger excess returns \cite{Fama1992TheCO}.
+Fama and French again innovated on the three-factor model in 2015 by adding two factors to the regression.
+In this model, Fama and French consider Investment and profitability while holding the other regression stable factors.
+The five-factor model analysis found that the addition of the investment and profitability factor made the value factor redundant \cite{Fama2013AFA}.
+The five-factor framework analysis has been recreated on international exchanges to test the effects on foreign markets.
+In recreation by Fama French and another study by Zhong and Li, the general relationships demonstrated in the five-factor model are also applicable in China, Japan, and Australian markets \cite{Fama2015InternationalTO, LIN2017141, Chiah2016A}.
+However, when the Fama French regression was applied to the Australian equity market, the authors did not find the significance of the value factor to subside like was visible in the 2015 paper \cite{Chiah2016A}.
+In addition to expanding the five-factor analysis geographically, some scholars have expanded the five-factor model by adding additional terms to assess momentum effects.
+In a paper published by  Dirkx and  Peter, the authors expanded the five-factor model by applying the Fama French framework to the German market while also adding a sixth term to assess the effect of risk premium with the CDAX constituents' momentum.
+Another study by Gregory and Stead used the sixth factor as an ESG factor to analyze if the ESG effect played a significant role in determining risk premia.
+This paper found that  "the risk premium of the sustainability factor is positive and significant\cite{Gregory2020TheGP}."
+
+In addition to Gregory and Stead's paper, several other authors have used the Fama French model to address the question concerning ESG and risk premia.
+To set up an ESG Fama French model, many academic journal articles use an index as a positive ESG portfolio and another index as a negative portfolio to create the ESG factor.
+In Gregory and Stead's paper, the author uses the S\&P ESG index as the high ESG index and the S\&P petroleum index as the low index to create a factor by subtracting the returns from the high ESG portfolio from the low ESG portfolio \cite{Gregory2020TheGP}.
+In Dorfleitner et al.'s paper, the authors collect Thomas Reuter's controversies and combined ESG score for 2500 companies.
+With the ranked ESG score data, the author then creates two portfolios by separating the top decile score and the bottom deciles scores, which act as the ESG and contra ESG portfolios.
+Similar to Gregory et al.
+paper, the author subtract the top 10\% from the bottom 10\% to create the ESG factor \cite{Dorfleitner2020ESGCA}.
 
@@
 An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index.
-Petroleum companies act as a stand-in for a contra ESG proflio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors.
+Petroleum companies act as a stand-in for a contra ESG portfolio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors.
 The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis.
@@
 The started was determined by the initiation of the MSCI ESG index.
-To collect daily individual stock return information CRSP data parsed from a larger dataset and then collapsed on month by summing each individual days returns to creat monthly returns.
+To collect daily individual stock return information CRSP data parsed from a larger data set and then collapsed on month by summing each individual days returns to create monthly returns.
 Similarly Fama French factors were collected on a daily basis and collapsed by month summing the daily returns for montly figures.
@@
 
-To capture abnormal returns we will consider 5 factors that is the 5 factors from the famma french paper small minus big(SMB),CMA, HML, and market portfolio(I.e model, consisting of market, size, value, profitability, and investment factors;).
-In addition I will ad another term to address how stocks move with the movement to a ESG porlfio.
-For this purpose we will use the MSCI ESG all world index.
 \section{Robustness Checks}
@@
 
+\section{Appendix}
 \vspace{1cm}
-\textbf{Table 2:} This table shows many things\\
+\textbf{Table 2:} This table shows the mean results by decile before any regressions\\
 \vspace{1cm}

### 2021-04-26 Paper/Main.tex
Updates from Overleaf

@@
 \begin{center}
-\input{Paper/prelim_beta}\\
+\input{Code/data/size_output_inital}
+\\
 \textbf{Table 5:} beta decile by estimated coefficients for the  five fama french factors and EMP factor
@@
 \bibliographystyle{plain}
-
+\singlespacing
 \newgeometry{margin=.15in}
-\input{Code/data/size_summary_stat_size}\\
 
 \section{Appendix}
-\vspace{1cm}
+\input{Code/data/size_summary_stat_size}\\
 \textbf{Table 2:} This table shows the mean results by decile before any regressions\\
 \vspace{1cm}
+
+\begin{center}
+\input{Code/data/beta_output_inital}\\
+\textbf{Table 3:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
+\end{center}
+\vspace{1cm}
+\begin{center}
+\input{Code/data/size_output_inital}\\
+\textbf{Table 4:} Size decile by estimated coefficients for the  five fama french factors and EMP factor\\
+\end{center}
+
+\begin{center}
+    \input{Code/data/FF_pass_2}\\
+    \textbf{Table 4:} second pass results for the size and beta decile portfolios\\
+\end{center}
+
+\begin{center}
+    \input{Code/data/FF_pass_2_no_ESG}\\
+    \textbf{Table 5:} second pass results for the size and beta decile portfolios without ESG factor\\
+\end{center}
 

### 2021-04-29 Paper/Main.tex
Updates from Overleaf

@@
 
-Global climate change presents a host of possible and sizable financial risks from increased water levels, which could threaten major global cities' safety, including Miami, Honk Konk, and Shanghai, with the potential to displace 100's of millions of people \cite{holder_three-degree_nodate}.
+Global climate change presents a host of possible and sizable financial risks from increased water levels, which could threaten major global cities' safety, including Miami, Honk Kong, and Shanghai, with the potential to displace 100's of millions of people \cite{holder_three-degree_nodate}.
 Global climate change is also shifting the landscape for global agricultural production as more extreme weather conditions, including droughts, pose problems to the most fundamental building block of economic activity.
@@
 In addition to risk investing professionals, ESG investing is appealing to a younger set of investors interested in a plethora of global issues from global climate change to diversity and inclusion in the workplace.
-A study conducted by Morgan Stanly found that millennials investors were twice as likely as the general investing population to invest in a company with social or environmental goals.
+A study conducted by Morgan Stanley found that millennials investors were twice as likely as the general investing population to invest in a company with social or environmental goals.
 
@@
 
- \begin{center}
-     \textbf{Table 1}\\
- \input{Code/data/desc1}
- \end{center}
-
 \section{Methodology}
@@
 \begin{split}
-    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t
+    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t +\epsilon_{i,t}
     \end{split}
@@
 \begin{equation}
-    E[R^e_{t}]= b_t\lambda_1+s_t\lambda_2 +h_t \lambda_3 +r_t\lambda_4 +c_t \lambda_4 + e_t \lambda_5
+    E[R^e]= b\lambda_1+s\lambda_2 +h\lambda_3 +r\lambda_4 +c \lambda_4 + e\lambda_5
     \label{second_pass}
 \end{equation}
-
-\begin{equation}
-    \frac{1}{T}\sum^T_{t=1}\lambda_{i,t}
-    \label{summing_lamda}
-\end{equation}
-\begin{center}
-    where:\\
-    T: is equal to the number of months in sample
-\end{center}
 
@@
 
-% size cofficents
-\begin{center}
-\input{Paper/prelim_size}\\
-\textbf{Table 3:} Size decile by estimated coefficients for the  five fama french factors and EMP factor
-\end{center}
-% size p vals
-\begin{center}
-\input{Paper/p_vals_size_prelim}\\
-\textbf{Table 4:} Size decile by p-values for table 3
-\end{center}
-% beta cofficents
-
 Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile.
@@
 
-\begin{center}
-\input{Code/data/size_output_inital}
-\\
-\textbf{Table 5:} beta decile by estimated coefficients for the  five fama french factors and EMP factor
-\end{center}
-% beta p vals
-\begin{center}
-\input{Paper/p_vals_beta_prelim}\\
-\textbf{Table 6:} Size decile by p-values for table 5
-\end{center}
-% pass 2
-\begin{center}
-\input{Paper/pass_2}\\
-\textbf{Table 7:} second pass results for the size and beta decile portfolios
-\end{center}
-
 \section{Robustness Checks}
+when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies 
  %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
@@
 \section{Appendix}
+
+ \begin{center}
+     \textbf{Table 1}\\
+ \input{Code/data/desc1}
+ \end{center}
 \input{Code/data/size_summary_stat_size}\\

### 2021-04-29 Paper/junk_maybe.tex

@@
 \end{equation}
+
+\begin{center}
+\input{Code/data/size_output_inital}
+\\
+\textbf{Table 5:} beta decile by estimated coefficients for the  five fama french factors and EMP factor
+\end{center}
+% beta p vals
+\begin{center}
+\input{Paper/p_vals_beta_prelim}\\
+\textbf{Table 6:} Size decile by p-values for table 5
+\end{center}
+% pass 2
+\begin{center}
+\input{Paper/pass_2}\\
+\textbf{Table 7:} second pass results for the size and beta decile portfolios
+\end{center}
+
+% size cofficents
+\begin{center}
+\input{Paper/prelim_size}\\
+\textbf{Table 3:} Size decile by estimated coefficients for the  five fama french factors and EMP factor
+\end{center}
+% size p vals
+\begin{center}
+\input{Paper/p_vals_size_prelim}\\
+\textbf{Table 4:} Size decile by p-values for table 3
+\end{center}
+% beta cofficents

### 2021-04-29 Paper/Main.tex
edit layout

@@
 \begin{document}
+
+%title page
+\input{Paper/title_page}
+
 \section{Introduction and Motivation}
-Asset pricing has a long history of models and theories governing how investors make decisions.
-Over the last decade, investors have become increasingly interested in factors outside of traditional financial metrics when assessing the possibility of an investment's returns.
-Much of this enthusiasm has been based on an ESG, Environmental Social Governance, framework.
-ESG investing considers how firms respond to these factors as potential impetuses for excess performance.
-This increased attention is especially visible in Morgan Stanley's 2019 survey of the ESG landscape.
-Morgan Stanley  found that 75\% of assessment managers say that "their firm has adopted sustainable investing." Only two years prior, in 2016, just 10\% of surveyed asset managed had adopted sustainable investing, demonstrating the rapid speed of adoption for ESG investing analysis \cite{morgan_stanley_sustainable_2019}.
-Practitioners have utilized the ESG framework to reduce risk in their investments while also finding abnormal returns.
-
-Global climate change presents a host of possible and sizable financial risks from increased water levels, which could threaten major global cities' safety, including Miami, Honk Kong, and Shanghai, with the potential to displace 100's of millions of people \cite{holder_three-degree_nodate}.
-Global climate change is also shifting the landscape for global agricultural production as more extreme weather conditions, including droughts, pose problems to the most fundamental building block of economic activity.
-In addition to risk mitigation, investors look to utilize ESG principle for superior returns.
-A research report published by McKinsey and Company linked gender diversity on executive teams to higher profitability and superior value creation \cite{mckinsey_diversity}.
-
-In addition to risk investing professionals, ESG investing is appealing to a younger set of investors interested in a plethora of global issues from global climate change to diversity and inclusion in the workplace.
-A study conducted by Morgan Stanley found that millennials investors were twice as likely as the general investing population to invest in a company with social or environmental goals.
-
-This new wave of investor sentiment has created strong tailwinds for sustainable investment projects.
-In 2020 alone, fund flows into sustainable funds grew by 51 billion dollars which is up over 100\% for the 2019 fund flow into sustainable investment products of about 21, placing total assets under management in sustainable products at over 250B.
-Fund flows into sustainable products for 2020 accounted for a little less a quarter of all US mutual fund flow and about one-fifteenth of all ETF fund flow according to morningstar \cite{monring_star_ESG}.
-
-To address the impact of ESG factors on stock preformance, this paper is broken into five sections.
-\textit{Section \ref{Literature}} will provide a  brief literature review of papers looking at abnormal performance around ESG criteria as well as discussion of asset pricing models to capture ESG factors in performance.
-Next \textit{ Section \ref{Data}}  will provide a brief discussion of the data used in this paper as well as the data cleaning process and summary statistics.
-Following the Data, \textit{Section \ref{Methodology}} will breakdown of the methodology for the paper.
-Finally, \textit{Section \ref{Results}}  will summarize the results of the regressions, \textit{Section \ref{Robustness}} will check the robustness of the results and \textit{Section \ref{Conclusion}} will provide a conclusion and possible extensions for next steps.
+\input{Paper/motivation}
 
@@
 \label{Literature}
-This section will be broke into two components.
-The first section will adress  theoretical models for asset purchasing and then we will introduce some of the mythology for  EGS factor in the asset pricing framework.\\
-
- \hl{add a section on the metholoogy for the 2 pass sort use} \cite{Kleibergen2013UnexplainedFA} and \cite{Fama1973RiskRA} \\
-The modern foundation of asset pricing began with the capital asset pricing model based on Sharp's 1964's work which used a regression analysis to relate market returns to individual security returns \cite{Sharpe1964CAPITALAP}.
-The CAPM model was later iterated on by Fama french in 1993 with the 3 factor, which considered several other terms to predict stock returns.
-Fama and French added size and book to market terms to the CAPM model.
-They found that these three coefficients created a more explanatory model which demonstrated the relationship that value companies(low book to market) and small companies experience larger excess returns \cite{Fama1992TheCO}.
-Fama and French again innovated on the three-factor model in 2015 by adding two factors to the regression.
-In this model, Fama and French consider Investment and profitability while holding the other regression stable factors.
-The five-factor model analysis found that the addition of the investment and profitability factor made the value factor redundant \cite{Fama2013AFA}.
-The five-factor framework analysis has been recreated on international exchanges to test the effects on foreign markets.
-In recreation by Fama French and another study by Zhong and Li, the general relationships demonstrated in the five-factor model are also applicable in China, Japan, and Australian markets \cite{Fama2015InternationalTO, LIN2017141, Chiah2016A}.
-However, when the Fama French regression was applied to the Australian equity market, the authors did not find the significance of the value factor to subside like was visible in the 2015 paper \cite{Chiah2016A}.
-In addition to expanding the five-factor analysis geographically, some scholars have expanded the five-factor model by adding additional terms to assess momentum effects.
-In a paper published by  Dirkx and  Peter, the authors expanded the five-factor model by applying the Fama French framework to the German market while also adding a sixth term to assess the effect of risk premium with the CDAX constituents' momentum.
-Another study by Gregory and Stead used the sixth factor as an ESG factor to analyze if the ESG effect played a significant role in determining risk premia.
-This paper found that  "the risk premium of the sustainability factor is positive and significant\cite{Gregory2020TheGP}."
-
-In addition to Gregory and Stead's paper, several other authors have used the Fama French model to address the question concerning ESG and risk premia.
-To set up an ESG Fama French model, many academic journal articles use an index as a positive ESG portfolio and another index as a negative portfolio to create the ESG factor.
-In Gregory and Stead's paper, the author uses the S\&P ESG index as the high ESG index and the S\&P petroleum index as the low index to create a factor by subtracting the returns from the high ESG portfolio from the low ESG portfolio \cite{Gregory2020TheGP}.
-In Dorfleitner et al.'s paper, the authors collect Thomas Reuter's controversies and combined ESG score for 2500 companies.
-With the ranked ESG score data, the author then creates two portfolios by separating the top decile score and the bottom deciles scores, which act as the ESG and contra ESG portfolios.
-Similar to Gregory et al.
-paper, the author subtract the top 10\% from the bottom 10\% to create the ESG factor \cite{Dorfleitner2020ESGCA}.
+\input{Paper/lit_review}
 
@@
 \label{Data}
-The data used for this paper was sourced from three datasets.
-The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices.
-Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily returns for the MSCI World ESG Focus Index and S\&P petroleum index, respectively.
-Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R.
-French - Data Library.
-
-An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index.
-Petroleum companies act as a stand-in for a contra ESG portfolio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors.
-The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis.
-In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}.
-Although the Petroleum industry has major environmental, risks there is less social orientated risks in petrolium.
-One solution to this provlem would be to use the index \$VICE which tracks alcohol, drug, gaming and betting industry  as the  contra ESG portfolio.
-However \$Vice has only traded for two years, as such the data is far less robust.
-Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor in Section \ref{Robustness}.
-It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry as has suffered from many human right violations which would significantly effects its social risk  \cite{Kapelus2002MiningCS}.
-
-\begin{equation}
-    \label{making_EMP}
-    EMP = R_{\text{World ESG Focus Index}}- R_{\text{S\&P petroleum index}}
-\end{equation}
-
-For this analysis, daily returns data was collected between June 2011 and March 2021.
-The started was determined by the initiation of the MSCI ESG index.
-To collect daily individual stock return information CRSP data parsed from a larger data set and then collapsed on month by summing each individual days returns to create monthly returns.
-Similarly Fama French factors were collected on a daily basis and collapsed by month summing the daily returns for montly figures.
-Finally the EMP factor was also  constructed summing the daily returns for each much.
-All three of the datasets were merged using the date and month as the matching component.
-The culmination of merging and collapsing resulted in  830,891 observations for monthly returns over  12,194 unique stocks.
-The merged and cleaned dataset had an average of 68 months of data of individual security data for each unique security.
-Below in table 1 are the summary statistics for the raw data.
+\input{Paper/data}
 
@@
 \label{Methodology}
-As discussed in the literature review, there are numerous models for assessing asset pricing and abnormal returns.
-In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model".
-This analysis method is built on the Fama French three factor model and relies on the two-pass sort with an assigned series of beta decile and size decile-based portfolios.
-Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors to assess relational interdependency.
-
-For this analysis, I sort each security into decile groups for beta and size (based on market cap).
-As firms' size and volatility are not constant functions, each firm is assigned a decile for each month depending on the most recent decile split.
-Each security will be assigned a decile for beta and size according to the CRSP /NYSE American beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size.
-After sorting each of the securities by beta and size decile, 300,264 of the original 830,889 data points had matches for both beta and size portfolios.
-To estimate the factor sensitivities base on the described decile groups, I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.
-
-\begin{equation}
-\begin{split}
-    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t +\epsilon_{i,t}
-    \end{split}
-    \label{six_factor_basic_reg}
-\end{equation}
-
-\begin{center}
-Where:\\
-    $R_{i,t}^e = R_{it}-R_{rf,t}$
-\end{center}
-
-Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort.
-The second pass of the two-pass sort is described in the equation  \eqref{second_pass} and is used to obtain the risk premia.
-To consolidate the risk premium across time, a simple average as described in equation \eqref{summing_lamda} is used to collapse the risk premia for each of the given factors.
-
-\begin{equation}
-    E[R^e]= b\lambda_1+s\lambda_2 +h\lambda_3 +r\lambda_4 +c \lambda_4 + e\lambda_5
-    \label{second_pass}
-\end{equation}
+\input{Paper/Methodology}
 
@@
 \label{Results}
-Table 3 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}.
-The column axis is the distinct size deciles where 1 is the smallest market cap companies, and 10 is the decline with the largest market cap companies.
-The row axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor.
-$HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.
-$SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks.
-$RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability.
-$CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms.
-$EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index.
-Finally, the intercept is the excess returns for the given decile.
-The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor.
-In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns.
-This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts.
-Table 4 complements table 3 with a breakdown of significance for each coefficient in table 3.
-The low p-values in the EMP factor indicate that the coefficients on all but the largest and smallest size deciles are significant at the 5\% level.
-It is important to note that all of the other Fama French factors are also less significant In the largest and small portfolios.
-This trend might indicate that the data might have pointed at far ends which do not behave like the other deciles.
-Table 6 demonstrates the risk premia for each of the factors using the second step in the two-pass sort.
-All of the Fama French factors in the size decile two-pass sort have positive risk premia.
-The EMP factor large than all of the other factors, with a coefficient of approximately  16.8\%
-
-Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile.
-In this table,  the beta decile 1 comprises the companies with the lowest beta.
-Similarly, beta decile 10 is the portfolio of companies with the largest beta values.
-Similar to the size decile, the EMP factor exhibits a negative relation with regard to its sensitivity.
-However, unlike the size decile analysis, there does not appear to be a significant relationship between the beta decile and the factor's sensitivity.
-Although the most significant impact is on the lowest decile, the least impact is experienced on the 6th decile portfolio, and the largest decile is close to the middle in terms of sensitivity.
-Similar to the size decile analysis, the EMP coefficients are all very significant, with every decile having significance at the 5\% level.
-When considering the second phase of the two-pass sort, the similarities to the size decile continue wherein each of the Fama French five factors exhibit positive risk premia, and the EMP also exhibits a positive relationship.
-The EMP factor again has a high coefficient with approximately 15.8\%.
+\input{Paper/Results}
 
 \section{Robustness Checks}
-when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies 
- %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
+\input{Paper/robustness}
 \label{Robustness}
@@
 \section{Appendix}
+%take these when needed
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/desc1}\\
+    \textbf{Table 1: EXPLAIN THE TABLE!!!}
+    \paperspacingwide
+\end{center}
 
 \begin{center}
-     \textbf{Table 1}\\
- \input{Code/data/desc1}
- \end{center}
+    \paperspacingnarrow 
     \input{Code/data/size_summary_stat_size}\\
     \textbf{Table 2:} This table shows the mean results by decile before any regressions\\
+    \paperspacingwide
+ \end{center}
 \vspace{1cm}
@@
 \begin{center}
+    \paperspacingnarrow
     \input{Code/data/beta_output_inital}\\
     \textbf{Table 3:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
+    \paperspacingwide
 \end{center}
+
 \vspace{1cm}
+
 \begin{center}
+    \paperspacingnarrow
     \input{Code/data/size_output_inital}\\
     \textbf{Table 4:} Size decile by estimated coefficients for the  five fama french factors and EMP factor\\
+    \paperspacingwide
 \end{center}
@@
 \begin{center}
+    \paperspacingnarrow
     \input{Code/data/FF_pass_2}\\
     \textbf{Table 4:} second pass results for the size and beta decile portfolios\\
+    \paperspacingwide
 \end{center}
@@
 \begin{center}
+    \paperspacingnarrow
     \input{Code/data/FF_pass_2_no_ESG}\\
     \textbf{Table 5:} second pass results for the size and beta decile portfolios without ESG factor\\
+    \paperspacingwide
+\end{center}
+
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/Industry_FF_pass_2}\\
+    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
+    \paperspacingwide
+
 \end{center}

### 2021-04-29 Paper/Methodology.tex
edit layout

As discussed in the literature review, there are numerous models for assessing asset pricing and abnormal returns. In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model". This analysis method is built on the Fama French three factor model and relies on the two-pass sort with an assigned series of beta decile and size decile-based portfolios. Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors to assess relational interdependency.

For this analysis, I sort each security into decile groups for beta and size (based on market cap). As firms' size and volatility are not constant functions, each firm is assigned a decile for each month depending on the most recent decile split.  Each security will be assigned a decile for beta and size according to the CRSP /NYSE American beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size. After sorting each of the securities by beta and size decile, 300,264 of the original 830,889 data points had matches for both beta and size portfolios. To estimate the factor sensitivities base on the described decile groups, I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.

\begin{equation}
\begin{split}
    R_{i,t}^e = a_i+b_i(R_{M,t}-R_{fr,t})+s_iSMB_t+ h_iHML_t+  r_iRMW_t+c_iCMA_t+e_iEMP_t +\epsilon_{i,t}
    \end{split}
    \label{six_factor_basic_reg}
\end{equation}

\begin{center}
Where:\\
    $R_{i,t}^e = R_{it}-R_{rf,t}$
\end{center}

Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort. The second pass of the two-pass sort is described in the equation  \eqref{second_pass} and is used to obtain the risk premia. To consolidate the risk premium across time, a simple average as described in equation \eqref{summing_lamda} is used to collapse the risk premia for each of the given factors.

\begin{equation}
    E[R^e]= b\lambda_1+s\lambda_2 +h\lambda_3 +r\lambda_4 +c \lambda_4 + e\lambda_5
    \label{second_pass}
\end{equation}

### 2021-04-29 Paper/Results.tex
edit layout

Table 3 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}. The column axis is the distinct size deciles where 1 is the smallest market cap companies, and 10 is the decline with the largest market cap companies. The row axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor. $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.  $SMB_{t}$ is the return on a diversified portfolio of small stocks minus the return on a diversified portfolio of big stocks. $RMW_t$ is the difference between the returns on diversified portfolios of stocks with robust and weak profitability. $CMA_{t}$ is the difference between the returns on diversified portfolios of the stocks of low and high investment firms. $EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index. Finally, the intercept is the excess returns for the given decile. The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor. In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns. This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts. Table 4 complements table 3 with a breakdown of significance for each coefficient in table 3. The low p-values in the EMP factor indicate that the coefficients on all but the largest and smallest size deciles are significant at the 5\% level. It is important to note that all of the other Fama French factors are also less significant In the largest and small portfolios. This trend might indicate that the data might have pointed at far ends which do not behave like the other deciles.  Table 6 demonstrates the risk premia for each of the factors using the second step in the two-pass sort. All of the Fama French factors in the size decile two-pass sort have positive risk premia. The EMP factor large than all of the other factors, with a coefficient of approximately  16.8\%

Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile. In this table,  the beta decile 1 comprises the companies with the lowest beta. Similarly, beta decile 10 is the portfolio of companies with the largest beta values.  Similar to the size decile, the EMP factor exhibits a negative relation with regard to its sensitivity. However, unlike the size decile analysis, there does not appear to be a significant relationship between the beta decile and the factor's sensitivity. Although the most significant impact is on the lowest decile, the least impact is experienced on the 6th decile portfolio, and the largest decile is close to the middle in terms of sensitivity. Similar to the size decile analysis, the EMP coefficients are all very significant, with every decile having significance at the 5\% level. When considering the second phase of the two-pass sort, the similarities to the size decile continue wherein each of the Fama French five factors exhibit positive risk premia, and the EMP also exhibits a positive relationship. The EMP factor again has a high coefficient with approximately 15.8\%.

### 2021-04-29 Paper/data.tex
edit layout

The data used for this paper was sourced from three datasets. The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices. Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily returns for the MSCI World ESG Focus Index and S\&P petroleum index, respectively. Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R. French - Data Library.

An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index. Petroleum companies act as a stand-in for a contra ESG portfolio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors. The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis. In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}. Although the Petroleum industry has major environmental, risks there is less social orientated risks in petrolium. One solution to this provlem would be to use the index \$VICE which tracks alcohol, drug, gaming and betting industry  as the  contra ESG portfolio. However \$Vice has only traded for two years, as such the data is far less robust. Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor in Section \ref{Robustness}. It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry as has suffered from many human right violations which would significantly effects its social risk  \cite{Kapelus2002MiningCS}.

\begin{equation}
    \label{making_EMP}
    EMP = R_{\text{World ESG Focus Index}}- R_{\text{S\&P petroleum index}}
\end{equation}

For this analysis, daily returns data was collected between June 2011 and March 2021. The started was determined by the initiation of the MSCI ESG index. To collect daily individual stock return information CRSP data parsed from a larger data set and then collapsed on month by summing each individual days returns to create monthly returns. Similarly Fama French factors were collected on a daily basis and collapsed by month summing the daily returns for montly figures. Finally the EMP factor was also  constructed summing the daily returns for each much. All three of the datasets were merged using the date and month as the matching component. The culmination of merging and collapsing resulted in  830,891 observations for monthly returns over  12,194 unique stocks. The merged and cleaned dataset had an average of 68 months of data of individual security data for each unique security. Below in table 1 are the summary statistics for the raw data.

### 2021-04-29 Paper/lit_review.tex
edit layout

This section will be broke into two components. The first section will adress  theoretical models for asset purchasing and then we will introduce some of the mythology for  EGS factor in the asset pricing framework.\\

 \hl{add a section on the metholoogy for the 2 pass sort use} \cite{Kleibergen2013UnexplainedFA} and \cite{Fama1973RiskRA} \\
The modern foundation of asset pricing began with the capital asset pricing model based on Sharp's 1964's work which used a regression analysis to relate market returns to individual security returns \cite{Sharpe1964CAPITALAP}. The CAPM model was later iterated on by Fama french in 1993 with the 3 factor, which considered several other terms to predict stock returns. Fama and French added size and book to market terms to the CAPM model. They found that these three coefficients created a more explanatory model which demonstrated the relationship that value companies(low book to market) and small companies experience larger excess returns \cite{Fama1992TheCO}. Fama and French again innovated on the three-factor model in 2015 by adding two factors to the regression. In this model, Fama and French consider Investment and profitability while holding the other regression stable factors. The five-factor model analysis found that the addition of the investment and profitability factor made the value factor redundant \cite{Fama2013AFA}. The five-factor framework analysis has been recreated on international exchanges to test the effects on foreign markets. In recreation by Fama French and another study by Zhong and Li, the general relationships demonstrated in the five-factor model are also applicable in China, Japan, and Australian markets \cite{Fama2015InternationalTO, LIN2017141, Chiah2016A}. However, when the Fama French regression was applied to the Australian equity market, the authors did not find the significance of the value factor to subside like was visible in the 2015 paper \cite{Chiah2016A}.  In addition to expanding the five-factor analysis geographically, some scholars have expanded the five-factor model by adding additional terms to assess momentum effects.  In a paper published by  Dirkx and  Peter, the authors expanded the five-factor model by applying the Fama French framework to the German market while also adding a sixth term to assess the effect of risk premium with the CDAX constituents' momentum. Another study by Gregory and Stead used the sixth factor as an ESG factor to analyze if the ESG effect played a significant role in determining risk premia. This paper found that  "the risk premium of the sustainability factor is positive and significant\cite{Gregory2020TheGP}."

In addition to Gregory and Stead's paper, several other authors have used the Fama French model to address the question concerning ESG and risk premia. To set up an ESG Fama French model, many academic journal articles use an index as a positive ESG portfolio and another index as a negative portfolio to create the ESG factor. In Gregory and Stead's paper, the author uses the S\&P ESG index as the high ESG index and the S\&P petroleum index as the low index to create a factor by subtracting the returns from the high ESG portfolio from the low ESG portfolio \cite{Gregory2020TheGP}. In Dorfleitner et al.'s paper, the authors collect Thomas Reuter's controversies and combined ESG score for 2500 companies. With the ranked ESG score data, the author then creates two portfolios by separating the top decile score and the bottom deciles scores, which act as the ESG and contra ESG portfolios. Similar to Gregory et al. paper, the author subtract the top 10\% from the bottom 10\% to create the ESG factor \cite{Dorfleitner2020ESGCA}.

### 2021-04-29 Paper/motivation.tex
edit layout

Asset pricing has a long history of models and theories governing how investors make decisions. Over the last decade, investors have become increasingly interested in factors outside of traditional financial metrics when assessing the possibility of an investment's returns. Much of this enthusiasm has been based on an ESG, Environmental Social Governance, framework. ESG investing considers how firms respond to these factors as potential impetuses for excess performance. This increased attention is especially visible in Morgan Stanley's 2019 survey of the ESG landscape. Morgan Stanley  found that 75\% of assessment managers say that "their firm has adopted sustainable investing." Only two years prior, in 2016, just 10\% of surveyed asset managed had adopted sustainable investing, demonstrating the rapid speed of adoption for ESG investing analysis \cite{morgan_stanley_sustainable_2019}. Practitioners have utilized the ESG framework to reduce risk in their investments while also finding abnormal returns.

Global climate change presents a host of possible and sizable financial risks from increased water levels, which could threaten major global cities' safety, including Miami, Honk Kong, and Shanghai, with the potential to displace 100's of millions of people \cite{holder_three-degree_nodate}. Global climate change is also shifting the landscape for global agricultural production as more extreme weather conditions, including droughts, pose problems to the most fundamental building block of economic activity. In addition to risk mitigation, investors look to utilize ESG principle for superior returns. A research report published by McKinsey and Company linked gender diversity on executive teams to higher profitability and superior value creation \cite{mckinsey_diversity}.

In addition to risk investing professionals, ESG investing is appealing to a younger set of investors interested in a plethora of global issues from global climate change to diversity and inclusion in the workplace. A study conducted by Morgan Stanley found that millennials investors were twice as likely as the general investing population to invest in a company with social or environmental goals.

This new wave of investor sentiment has created strong tailwinds for sustainable investment projects. In 2020 alone, fund flows into sustainable funds grew by 51 billion dollars which is up over 100\% for the 2019 fund flow into sustainable investment products of about 21, placing total assets under management in sustainable products at over 250B. Fund flows into sustainable products for 2020 accounted for a little less a quarter of all US mutual fund flow and about one-fifteenth of all ETF fund flow according to morningstar \cite{monring_star_ESG}.

To address the impact of ESG factors on stock preformance, this paper is broken into five sections. \textit{Section \ref{Literature}} will provide a  brief literature review of papers looking at abnormal performance around ESG criteria as well as discussion of asset pricing models to capture ESG factors in performance. Next \textit{ Section \ref{Data}}  will provide a brief discussion of the data used in this paper as well as the data cleaning process and summary statistics. Following the Data, \textit{Section \ref{Methodology}} will breakdown of the methodology for the paper. Finally, \textit{Section \ref{Results}}  will summarize the results of the regressions, \textit{Section \ref{Robustness}} will check the robustness of the results and \textit{Section \ref{Conclusion}} will provide a conclusion and possible extensions for next steps.

### 2021-04-29 Paper/robustness.tex
edit layout

when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies 
 %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]

### 2021-04-29 Paper/title_page.tex
edit layout

\begin{titlepage}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

\center % Center everything on the page

%----------------------------------------------------------------------------------------
%	HEADING SECTIONS
%----------------------------------------------------------------------------------------

\textsc{\LARGE Colgate University}\\[1.5cm] % Name of your university/college
\includegraphics[scale=.1]{Paper/secondary/colgate_logo.png}\\[1cm] % Include a department/university logo - this will require the graphicx package
\textsc{\Large Financial Economics Seminar}\\[0.5cm] % Major heading such as course name
\textsc{\large Econ 487}\\[0.5cm] % Minor heading such as course title

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\HRule \\[0.4cm]
{ \huge \bfseries Factor Investing with EGS}\\[0.4cm] % Title of your document
\HRule \\[1.5cm]

%----------------------------------------------------------------------------------------
%	AUTHOR SECTION
%----------------------------------------------------------------------------------------

\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}
Matthew \textsc{Chistolini}\\ % Your name
\end{flushleft}

\end{minipage}\\[1cm]

% If you don't want a supervisor, uncomment the two lines below and remove the section above
\Large \emph{Advisor:}
Michael \textsc{Connolly}\\[2cm] % Your name

%----------------------------------------------------------------------------------------
%	DATE SECTION
%----------------------------------------------------------------------------------------

{\large \today}\\[2cm] % Date, change the \today to a set date if you want to be precise

\vfill % Fill the rest of the page with whitespace

\end{titlepage}

### 2021-05-02 Paper/Methodology.tex
paused work for April 28
big push in latex

@@
 In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model".
-This analysis method is built on the Fama French three factor model and relies on the two-pass sort with an assigned series of beta decile and size decile-based portfolios.
-Breaking the portfolios into these distinct groups allows for further analysis of the effect across these two factors to assess relational interdependency.
+This analysis method is built on the Fama French three-factor model and relies on the two-pass sort.
+Like the section \ref{Data} on data, this section is broken up into multiple parts.
+First, I will explain the methodology for assessing the sensitivity and risk premium for the ESG factor created from the MSCI ESG index minus the petroleum index.
+Next, I will discuss the methodology for MSCI ESG rating ESG factors.
+I will conclude with a brief description of the newey west errors and the setup of assessing statistical significance.
 
-For this analysis, I sort each security into decile groups for beta and size (based on market cap).
-As firms' size and volatility are not constant functions, each firm is assigned a decile for each month depending on the most recent decile split.
-Each security will be assigned a decile for beta and size according to the CRSP /NYSE American beta decile for beta and the CRSP Nasdaq Market Capitalization portfolio for size.
-After sorting each of the securities by beta and size decile, 300,264 of the original 830,889 data points had matches for both beta and size portfolios.
-To estimate the factor sensitivities base on the described decile groups, I ran the regression in equation \eqref{six_factor_basic_reg} on the cleaned data set.
+For the first part of this study, I sort each security into decile groups based on beta and size.
+The size and beta deciles ratings are provided by CRSP and are merged on the individual firm level for each month of data.
+Next, I conducted the first pass for the two-pass sort by running the regression of expected returns on market risk premia the four Fama French factors and the EMP factor as described in equation \eqref{six_factor_basic_reg}.
+This regression was run twice: once on the beta-ranked portfolios and again on the size-ranked portfolios.
 
@@
 Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort.
-The second pass of the two-pass sort is described in the equation  \eqref{second_pass} and is used to obtain the risk premia.
-To consolidate the risk premium across time, a simple average as described in equation \eqref{summing_lamda} is used to collapse the risk premia for each of the given factors.
+The second pass of the two-pass sort is described in the equation  \eqref{second_pass}.
+The second pass describes the risk premia for each of the six factors in the mode.
+This second pass was performed for both beta and size-based portfolios.
 
@@
     \label{second_pass}
-\end{equation}
+\end{equation}\\
+
+The second set of regressions is focused on establishing a link between MSCI ESG ratings and ESG risk factor sensitivity.
+Similar to the first set of regressions, I use a two-pass sort methodology using both size and beta ranked portfolios.
+To create the ESG factor for this set of analyses, I used the all industry ESG factor described in Section \ref{Data}.
+The process of implementing the regression follows the same basic form as the one for the MSCI ESG index and petroleum index ESG factor by first running a pass on the portfolio by size and following that with a cross-sectional analysis to quantify the risk premium.
+
+Finally, in determining the statistical significance of each of the coefficients, it is important to discuss the newly west error.
+The newly west error is a replacement for standard error in a traditional OLS regression.
+the advantage of the newly west is an ability to account for serial correlation within the model

### 2021-05-02 Paper/Results.tex
big push in latex

@@
-Table 3 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}.
-The column axis is the distinct size deciles where 1 is the smallest market cap companies, and 10 is the decline with the largest market cap companies.
-The row axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor.
+Table 2 represents a matrix of values for the first pass in the Fama French two-pass sort, described in equation \eqref{six_factor_basic_reg}.
+The row axis is the distinct size deciles, where 1 is the smallest 10\% of companies by market cap companies, and 10 is the decile with the largest 10\% of companies by market cap.
+The column axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor.
 $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.
@@
 In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns.
-This relationship indicates that investors are actually paying a premium to invest in ESG companies over the contra ESG counterparts.
-Table 4 complements table 3 with a breakdown of significance for each coefficient in table 3.
-The low p-values in the EMP factor indicate that the coefficients on all but the largest and smallest size deciles are significant at the 5\% level.
-It is important to note that all of the other Fama French factors are also less significant In the largest and small portfolios.
-This trend might indicate that the data might have pointed at far ends which do not behave like the other deciles.
-Table 6 demonstrates the risk premia for each of the factors using the second step in the two-pass sort.
-All of the Fama French factors in the size decile two-pass sort have positive risk premia.
-The EMP factor large than all of the other factors, with a coefficient of approximately  16.8\%
+This relationship indicates that investors are returning a premium by investing in ESG companies over the contra ESG counterparts and that this relationship is largest for small companies.
+The results from this statistical test are significant at the 5\% level for the EMP factor for all of the portfolios except the smallest and largest size portfolios.
+However, several of the Fama French factors are also not significant for the largest and smallest portfolios, which may indicate that these tail portfolios might be behaving differently due to their composition rather than a reflection of the factor.
 
-Table 5 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the column axis is the beta decile.
-In this table,  the beta decile 1 comprises the companies with the lowest beta.
-Similarly, beta decile 10 is the portfolio of companies with the largest beta values.
-Similar to the size decile, the EMP factor exhibits a negative relation with regard to its sensitivity.
-However, unlike the size decile analysis, there does not appear to be a significant relationship between the beta decile and the factor's sensitivity.
-Although the most significant impact is on the lowest decile, the least impact is experienced on the 6th decile portfolio, and the largest decile is close to the middle in terms of sensitivity.
-Similar to the size decile analysis, the EMP coefficients are all very significant, with every decile having significance at the 5\% level.
-When considering the second phase of the two-pass sort, the similarities to the size decile continue wherein each of the Fama French five factors exhibit positive risk premia, and the EMP also exhibits a positive relationship.
-The EMP factor again has a high coefficient with approximately 15.8\%.
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/size_output_inital}\\
+    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size.
+the row axis is portfolio and the column axis is the  market beta,four Fama French factors and the EMP factor\\
+    \paperspacingwide
+\end{center}
+
+Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for Size decile portfolios.
+The market factor, the SMB factor, the CMA factor, and the EMP factor are all positive and statistically significant at the 1\% level.
+The other two factors have a negative sign and are not statistically significant at the 10\% level.
+EMP factor demonstrates that a 1 unit change in beta corresponds with an 8.7\% increase in returns.
+Although this is by far the smallest coefficient of the four statistically significant returns, this term is still economically meaningful and represents a risk premia for investing in high ESG firms.
+Overall the model appears to be a fairly good fit with a $R^2$ value of 87.5\%, which is lower than some previous literature using the petroleum index but still indicates a well fit model \hl{cite this paper}
+
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/FF_pass_2}\\
+    \textbf{Table 3:} second pass results for the size and beta decile portfolios\\
+    \paperspacingwide
+\end{center}
+
+Table 4 details the relationship outlined in equation \eqref{six_factor_basic_reg} in the methodology where the row axis is the beta decile.
+In this table,  the beta decile 1 comprises the companies with the lowest market beta's.
+Similarly, beta decile 10 is the portfolio of companies with the largest market beta values.
+Unlike the size decile portfolios, the beta decile portfolios do not have consistently negative relations with respect to decile.
+Rather the relationship appears to have a dome shape where portfolio 6 has the smallest positive sensitivity to the EMP factor, and larger or smaller beta portfolios appear to have a larger sensitivity to the  EMP factor.
+\hl{why might this be}.
+Similar to the size decile analysis, the EMP coefficients are all very significant, with every portfolio having significance at the 5\% level.
+
+When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB CMA and EMP exhibit positive risk premia.
+The other two factors remain statistically insignificant, and the MHL factor changed signs and now has a positive relationship.
+The EMP coefficient indicates that a one-unit change in beta would result in a 7.36\% change in returns for a high ESG company.
+This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude.
+The model is also a good fit with an $R^2$ value of 86.0%
+
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/beta_output_inital}\\
+    \textbf{Table 4:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
+    \paperspacingwide
+\end{center}
+
+Moving on to the MSCI based portfolio, a two-pass sort was also used to assess the sensitivity and risk premia on the ESG portfolio on both size and  beta porflio
+
+\begin{center}
+    \paperspacingnarrow
+\input{Code/data/All_Industry_FF_pass_2}\\
+\textbf{Table 7:} SIC top 10\% minus bottom 10\%
+    \paperspacingwide
+\end{center}

### 2021-05-02 Paper/data.tex
big push in latex

@@
-The data used for this paper was sourced from three datasets.
+The data used for this paper was sourced from four datasets.
 The daily performance of the individual stock prices was collected from CRSP's primarily daily stock prices.
-Data on the ESG portfolio and the contra ESG portfolio were extracted from Compustat as the daily returns for the MSCI World ESG Focus Index and S\&P petroleum index, respectively.
+Data for the daily returns of MSCI World ESG Focus Index and S\&P petroleum index was sourced from CAPIQ Compustat.
+Data on firm-level ESG scores (by ticker) was sourced from MSCI and reflected the most recent company ratings as MSCI does not have the historical ratings relatively available.
 Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R.
@@
 
-An ESG factor, denoted as EMP (Esg Minus Petroleum), was created by taking the daily returns from the MSCI World ESG Focus Index and subtracting the daily returns for the the  S\&P petroleum index.
-Petroleum companies act as a stand-in for a contra ESG portfolio in a similar way to isolate the specific risk component just to ESG similar to how Fama frech create their factors.
-The  petroleum index was chosen as non renewable energy industry is a exposed to a number of environmental risk factors that are considered substantially in an ESG analysis.
-In addition, the use of a petroleum index has been used in other academic work as a stand in for a contra ESG portfolio \cite{Frynas2005TheFD}.
-Although the Petroleum industry has major environmental, risks there is less social orientated risks in petrolium.
-One solution to this provlem would be to use the index \$VICE which tracks alcohol, drug, gaming and betting industry  as the  contra ESG portfolio.
-However \$Vice has only traded for two years, as such the data is far less robust.
-Instead, to check for robustness of the ESG factor and to exposure the contra portfolio to both strong social and environmental risk factors the analysis explained in the methodology will be repeated using the \& Mining Select Industry Index as the contra ESG portfolio for the EMP factor in Section \ref{Robustness}.
-It has been also well documented that the mining industry also struggles to adapt to ESG factors due to the nature of the industry as has suffered from many human right violations which would significantly effects its social risk  \cite{Kapelus2002MiningCS}.
+The data and methodology for the paper are broken into two sections; first, a preliminary study is conducted using a broad base ESG factor to confirm a statistically risk sensitivity relationship between ESG and returns.
+Next, an analysis on a SIC level is conducted in an attempt to understand how industry-specific ESG factors affect the risk-return relationship of stock performance.
 
+In the first stage of the analysis, an ESG factor is created by taking two market indexes and subtracting their daily returns in an attempt to isolate ESG based risk.
+The formula for this factor is represented in Equation \eqref{making_EMP}, and for the remainder of the section, one analysis will be referred to as EMP, which is sort for ESG minus petroleum.
+This method of creating a factor is very similar to Fama and French in their other four factors in the five-factor model.
+For this analysis, I will be using the MSCI World ESG Focus Index as the high ESG portfolio and the  S\&P petroleum index as the low ESG portfolio.
+The methodology for creating an ESG factor is similar to the work from \hl{put the author's name in here}.
+In \hl{put the author's name in here}'s work, they were able to find a statistical risk relationship between a high ESG portfolio and a petroleum index\cite{Frynas2005TheFD}.
+The petroleum index provides a good contra ESG portfolio as the petroleum industry is in direct opposition to environmental goals that are E components of ESG.
+A similar analysis could also be run with other market indexes.
+The S\&P Mining index is another popular index that also performs very poorly for environmental and social concerns\cite{Kapelus2002MiningCS}.
+A final index that also could represent a contra ESG portfolio is the  \$VICE ETF.
+This financial product tracks the market for  alcohol, drug, gaming, and betting industry which have all been shown as clear contra ESG industries \hl{inset citation}
 \begin{equation}
@@
 
-For this analysis, daily returns data was collected between June 2011 and March 2021.
-The started was determined by the initiation of the MSCI ESG index.
-To collect daily individual stock return information CRSP data parsed from a larger data set and then collapsed on month by summing each individual days returns to create monthly returns.
-Similarly Fama French factors were collected on a daily basis and collapsed by month summing the daily returns for montly figures.
-Finally the EMP factor was also  constructed summing the daily returns for each much.
-All three of the datasets were merged using the date and month as the matching component.
-The culmination of merging and collapsing resulted in  830,891 observations for monthly returns over  12,194 unique stocks.
-The merged and cleaned dataset had an average of 68 months of data of individual security data for each unique security.
-Below in table 1 are the summary statistics for the raw data.
+To conduct this analysis, daily data was collected between June 2011 and March 2021.
+June 2011 marks the initiation of the MSCI ESG index; as such, this is the beginning of the sample period for this paper.
+Daily individual stock return information was collected from CRSP and merged with the daily ESG factor returns.
+The return data was collapsed on month to create a monthly return figure by summing all of the daily returns for each given month in the dataset.
+Next, the Fama French monthly factors were merged into the dataset for each individual firm for each month.
+After aggregating, the factor and individual firm return, there were 830,891 observations of monthly return data for 12,194 unique stocks.
+The merged and cleaned dataset had an average of 68 months of data for each individual security.
+Table 1 provides brief summary statistics for the data in the first stage of the analysis.
+
+In the second stage of the analysis, I look at how SIC code-specific ESG factors affect the risk-return relationship.
+In this study, I used the MSCI ESG rating by ticker to created SIC-based ESG factors.
+MSCI rate companies on their ESG score AAA-CCC where AAA is the highest ESG rating and CCC is the lowest ESG rating.
+In the sample of MSCI ratings, 16 companies were AAA, 88 companies were AA, 118 companies were A, 170 companies were BBB, 108 companies were BB, 40 companies were B, and 5 companies earned a CCC rating.
+Each of these ratings was assigned a mating numeric score (1-7) and were merged using the CRSP name file to their corresponding kypermno value by matching ticker information.
+Subsequently, the ESG scores were merged into the monthly return dataset described at the end of the first analysis.
+A final process was conducted to make a unique ESG factor for each of the SIC codes.
+The unique ESG factor was created at the first digit SIC level and generated by splitting the SIC level ESG data into deciles and taking the top 10\% and as the high ESG index bottom 10\% as the low ESG and subtracting high ESG minus low ESG as explained in equation \eqref{making_EMP_section}.
+This ESG factor was created for SIC codes beginning with 1-4 and 6-8.
+SIC code.
+SIC code 5 only had BBB ratings, and MSCI did not have significant data on SIC code 0 or 9.
+In addition to creating industry specific ESG factors I also created a all industry ESG factor using the top 10\% and bottom 10\% for all MSCI ESG scores to create an ESG factor.
+\hl{explain how EMP is the same down here for consistency of reuslts}
+
+\begin{equation}
+    \label{making_EMP_section}
+    EMP_i = R_{\text{high ESG,i}}- R_{\text{low ESG,i}}
+\end{equation}
+
+\begin{center}
+    Where i is the first digit SIC code    
+\end{center}

### 2021-05-02 Paper/Results.tex
latex update

@@
 
-Moving on to the MSCI based portfolio, a two-pass sort was also used to assess the sensitivity and risk premia on the ESG portfolio on both size and  beta porflio
+Moving on to the second stage of the analysis, I constructed an ESG factor out of the MSCI ESG ratings to capture a more holistic view of ESG risk.
+I created this factor by taking the top 10\% of MSCI EGS rated companies and subtracted them from the bottom 10\% of firms rated by the MSCI.
+I repeated the methodology above for the regression analysis.
+I  regressed two sets of two-pass regression analyses one on beta-sized portfolios and another on size-based portfolios.
+In the first pass results from the new factor, the beta portfolio demonstrated a decrease in sensitivity as the betas grew; however, the size-based portfolios did not experience any real pattern governing the decile and the sensitivity.
+In the case of the size-based portfolios, some of the sensitivity figures were positive while others were negative.
+Table 4 represents a second stage two-pass sort regression output from this table it is clear that
 
@@
 \input{Code/data/All_Industry_FF_pass_2}\\
-\textbf{Table 7:} SIC top 10\% minus bottom 10\%
+\textbf{Table 4:} SIC top 10\% minus bottom 10\%
     \paperspacingwide

### 2021-05-02 Paper/robustness.tex
big push in latex
latex update

@@
  %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
+
+ dicuss limitation of only one set of ESG factors for MSCI
+
+ we used the Size deciles on a go froward basis when analysis the data as the statistical relationship was more apparent
+
+ \textbf{Analysis for the w/o ESG on the data}\\
+ \textbf{yearly effect}\\
+
+ \textbf{MSCI data 2019 beyond}\\
+
+ \textbf{MSCI SIC level analysis}\\
+
+ problems we can never solve for both of these indexes does not do a great job capturing small companies so size might be a bit of a probelm overall as the small companies do not have ther ESG scores baked in as MSCI does not have the resources to rank every company

### 2021-05-02 Paper/Main.tex
paused work for April 28
big push in latex
Updates from Overleaf

@@
     \paperspacingnarrow
-    \input{Code/data/beta_output_inital}\\
-    \textbf{Table 3:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
-    \paperspacingwide
-\end{center}
-
-\vspace{1cm}
-
-\begin{center}
-    \paperspacingnarrow
-    \input{Code/data/size_output_inital}\\
-    \textbf{Table 4:} Size decile by estimated coefficients for the  five fama french factors and EMP factor\\
-    \paperspacingwide
-\end{center}
-
-\begin{center}
-    \paperspacingnarrow
     \input{Code/data/FF_pass_2}\\
@@
     \paperspacingwide
-
 \end{center}
 
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/Time_FF_pass_2}\\
+    \textbf{Table 7:} Time based cofficents
+    \paperspacingwide
+\end{center}
+
+\begin{center}
+    \paperspacingnarrow
+\input{Code/data/All_Industry_FF_pass_2}\\
+\textbf{Table 7:} SIC top 10\% minus bottom 10\%
+    \paperspacingwide
+\end{center}
+
+\begin{center}
+    \paperspacingnarrow
+\input{Code/data/All_Industry_FF_pass_2_past_2018}\\
+\textbf{Table XX:} SIC top 10\% minus bottom 10\%
+    \paperspacingwide
+\end{center}
 \clearpage

### 2021-05-02 Paper/Main.tex
Updates from Overleaf

@@
 %take these when needed
-\begin{center}
-    \paperspacingnarrow
-    \input{Code/data/desc1}\\
-    \textbf{Table 1: EXPLAIN THE TABLE!!!}
-    \paperspacingwide
-\end{center}
-
-\begin{center}
-    \paperspacingnarrow 
-    \input{Code/data/size_summary_stat_size}\\
-    \textbf{Table 2:} This table shows the mean results by decile before any regressions\\
-    \paperspacingwide
- \end{center}
-\vspace{1cm}
-
-\begin{center}
-    \paperspacingnarrow
-    \input{Code/data/FF_pass_2}\\
-    \textbf{Table 4:} second pass results for the size and beta decile portfolios\\
-    \paperspacingwide
-\end{center}
 
@@
 \input{Code/data/All_Industry_FF_pass_2_past_2018}\\
-\textbf{Table XX:} SIC top 10\% minus bottom 10\%
+\textbf{Table XX:} SIC top 10\% minus bottom 10\% 2018-2020
     \paperspacingwide

### 2021-05-02 Paper/Results.tex
Updates from Overleaf

@@
 
-When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB CMA and EMP exhibit positive risk premia.
+When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB, CMA, and EMP exhibit positive risk premia.
 The other two factors remain statistically insignificant, and the MHL factor changed signs and now has a positive relationship.
@@
 This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude.
-The model is also a good fit with an $R^2$ value of 86.0%
+The model is also a good fit with an $R^2$ value of 86.0\%.
+With the baseline analysis proving that there is a relationship between ESG and excess marker returns, I will provide an investigation of the robustness in section \ref{Robustness} by first comparing the results of the two-pass sort with the EGS factor to the same analysis without the factor.
+Next, I will investigate time-related relationships between ESG and excess returns by implementing a dummy variable for the year and repeating the two-pass sort methodology.
 
@@
 I repeated the methodology above for the regression analysis.
-I  regressed two sets of two-pass regression analyses one on beta-sized portfolios and another on size-based portfolios.
-In the first pass results from the new factor, the beta portfolio demonstrated a decrease in sensitivity as the betas grew; however, the size-based portfolios did not experience any real pattern governing the decile and the sensitivity.
+I  regressed two sets of two-pass regression analyses, one on beta-sized portfolios and another on size-based portfolios.
+In the first pass results for the MSCI factor, the beta portfolio demonstrated a decrease in sensitivity as the betas grew; however, the size-based portfolios did not experience any real pattern governing the decile and the sensitivity.
 In the case of the size-based portfolios, some of the sensitivity figures were positive while others were negative.
-Table 4 represents a second stage two-pass sort regression output from this table it is clear that
+Table 5 represents a second stage two-pass sort regression output.
+Similar to the regression in the first stage of the analysis, the market return, CMA, SMB remain statistically significant; however, only the size-based portfolios demonstrate a statistically significant coefficient for the ESG factor.
+The EMP ESG factor coefficient can be interpreted as a negative relationship between ESG and excess return wherein a one beta unit increase in ESG results in a decrease of 2\% in excess returns.
+One limitation of the MSCI data is it only represents a snapshot in ESG rating data; that is, the MSCI data is only the most recent ESG performance and does not account for changes in ESG performance.
+As such, a company that used to be a  poor ESG and has improved to a strong ESG company will only be categorized as a strong ESG company which will not accurately address the ESG performance.
+As such, in section \ref{Robustness} I will address this by shrinking the time frame to see if the coefficients increase in their significance.
+Further,  to address additional concerns of robustness, I will run another two-pass sort to isolate the SIC industry-specific ESG factor and determine if there are ESG related excess returns that are greater in some industries than others.
 
@@
     \input{Code/data/All_Industry_FF_pass_2}\\
-\textbf{Table 4:} SIC top 10\% minus bottom 10\%
+    \textbf{Table 5:} SIC top 10\% minus bottom 10\%
     \paperspacingwide

### 2021-05-03 Paper/Main.tex
latex now basically to the end!!

@@
 \label{Conclusion}
+\input{Paper/conclusion}
 
@@
 \bibliographystyle{plain}
-\singlespacing
-\newgeometry{margin=.15in}
-
-\section{Appendix}
-%take these when needed
-
-\begin{center}
-    \paperspacingnarrow
-    \input{Code/data/FF_pass_2_no_ESG}\\
-    \textbf{Table 5:} second pass results for the size and beta decile portfolios without ESG factor\\
-    \paperspacingwide
-\end{center}
-
-\begin{center}
-    \paperspacingnarrow
-    \input{Code/data/Industry_FF_pass_2}\\
-    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
-    \paperspacingwide
-\end{center}
-
-\begin{center}
-    \paperspacingnarrow
-    \input{Code/data/Time_FF_pass_2}\\
-    \textbf{Table 7:} Time based cofficents
-    \paperspacingwide
-\end{center}
-
-\begin{center}
-    \paperspacingnarrow
-\input{Code/data/All_Industry_FF_pass_2}\\
-\textbf{Table 7:} SIC top 10\% minus bottom 10\%
-    \paperspacingwide
-\end{center}
-
-\begin{center}
-    \paperspacingnarrow
-\input{Code/data/All_Industry_FF_pass_2_past_2018}\\
-\textbf{Table XX:} SIC top 10\% minus bottom 10\% 2018-2020
-    \paperspacingwide
-\end{center}
-\clearpage
-\restoregeometry
 

### 2021-05-03 Paper/conclusion.tex
latex now basically to the end!!

Looking beyond this paper's scope and future research opportunities, there are a few natural extensions for assessing the role rating agencies and indexes play in the EGS landscape. One extension is to further isolate ESG premia by conducting an event study methodology on several well-known ESG indexes, including the MSCI ESG leaders index referenced in this paper. Using addition and deletion events from this index can provide a tangible event to test the market reaction to changes in ESG news and assess if there are abnormal returns. This methodology for evaluating abnormal returns can also be applied to ESG rating; however, there might be less of an effect if the market was anticipating a change in the rating. As such, the indexing market event might provide a more precise image regarding market tracking behavior around ESG.

Another possible extension for ESG research is to take a further look at the rating agencies. Over the last few years, more companies have entered into the ESG rating space, MSCI and Sustainalytics have held a dominant position in the market, but S\&P, Moody's, and other firms are looking to expand their research efforts around ESG research. An interesting extension would be to analyze the risk premia for different firms and see if there an ESG factor where high ESG minus low ESG has a larger risk premium for one rating agency over another rating agency.

### 2021-05-03 Paper/Results.tex
latex now basically to the end!!
Updates from Overleaf

@@
     \input{Code/data/FF_pass_2}\\
-    \textbf{Table 3:} second pass results for the size and beta decile portfolios\\
+    \textbf{Table 3:} shows the results for the Second pass of the twp pass sort.
+On the left labeled as (1) are the regression results for the size decile portfolios.
+On the right labeled as (2) are the regression results for the beta decile portfolios\\
     \paperspacingwide
@@
     \input{Code/data/All_Industry_FF_pass_2}\\
-    \textbf{Table 5:} SIC top 10\% minus bottom 10\%
+    \textbf{Table 5:} SIC top 10\% minus bottom 10\% \hl{NEED TO EDIT THESE RESULTS}
     \paperspacingwide

### 2021-05-03 Paper/robustness.tex
Updates from Overleaf
latex now basically to the end!!

@@
+As discussed in section \ref{Results}, additional tests are required to ensure the methodology's robustness.
+The first robustness check I will run is a repetition of the two-pass sort for where the ESG factor is calculated asMSCI ESG index minus the petroleum index (EMP).
+Table 6 confirms the statistical significance of the market term, the SMB term, and the CMA term.
+All three of these terms were found to be statistically significant in the initial regression as well.
+Beyond the coefficients,  the results in table 6 demonstrate a slightly lower $R^2$ for both the Beta decile and the size decile variants, demonstrating that the EMP term contributes to the explanation of excess returns.
+In addition, the constant term for both size and beta decile portfolios increased in magnitude as a result of omitting the EMP.
+As the coefficient term gets closer to zero, the factors do a better job describing the returns and indicate that there is no excess return that is not explained.
+ %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
+
+ \begin{center}
+    \paperspacingnarrow
+    \input{Code/data/FF_pass_2_no_ESG}\\
+    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
+    \paperspacingwide
+\end{center}
+Another robustness check that is essential to consider is the time-based effects on the model.
+Although we have demonstrated that the EMP factor is statistically significant and explains part of our model, it is essential to understand if this is ubiquitous across time in the sample or if the  ESG factor has shifted in risk premia over time.
+To address this robustness concern, I have repeated the two-pass sort with the size decile portfolio by adding a time-based dummy variable.
+Table 7 details the results from the regressions.
+In 2015 2016, 2018, and 2019 there is statistical significance for the EMP term and a positive association between EMP and excess returns.
+The ESG factor showed the most strength in 2019 as well as the greatest coefficient value.
+In 2019 a one-unit change in beta would result in a 25.9\% increase in excess returns.
+The coefficient values in 2017 appeared to hold a negative relationship as a one-unit change in ESG beta would have resulted in a reduction in excess returns by 10.5\%.
+Although in 2020, there was a positive relationship, the coefficient is not statistically significant.
+However, in 2020, the CMA factor, which has been statistically significant at the 1\% level for all other tests, also does not appear to be statistically significant.
+2020 represents a bit of a deviation due to the larger market condition from the COVID-19 pandemic rather than the EMP factor's long-run behavior.
+
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/Time_FF_pass_2}\\
+    \textbf{Table 7:} Shows the results for the second pass in a two pass sort. The two pass sort used size based deciles for the first pass. Columns (1)-(6) represent individual years of data. 
+    \paperspacingwide
+\end{center}
+ \textbf{MSCI data 2018-2020}\\
+ \begin{center}
+    \paperspacingnarrow
+\input{Code/data/All_Industry_FF_pass_2_past_2018}\\
+\textbf{Table 8:} SIC top 10\% minus bottom 10\% 2018-2020
+    \paperspacingwide
+\end{center}
+ \textbf{MSCI SIC level analysis}\\
+
+  \begin{center}
+    \paperspacingnarrow
+    \input{Code/data/Industry_FF_pass_2}\\
+    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
+    \paperspacingwide
+\end{center}
+
+ WE ADJUSTED THE YEAR to 2019!!
+ problems we can never solve for both of these indexes does not do a great job capturing small companies so size might be a bit of a probelm overall as the small companies do not have ther ESG scores baked in as MSCI does not have the resources to rank every company
+
  when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies
- %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
 
@@
  we used the Size deciles on a go froward basis when analysis the data as the statistical relationship was more apparent
-
- \textbf{Analysis for the w/o ESG on the data}\\
- \textbf{yearly effect}\\
-
- \textbf{MSCI data 2019 beyond}\\
-
- \textbf{MSCI SIC level analysis}\\
-
- problems we can never solve for both of these indexes does not do a great job capturing small companies so size might be a bit of a probelm overall as the small companies do not have ther ESG scores baked in as MSCI does not have the resources to rank every company

### 2021-05-03 Paper/Results.tex

@@
     \input{Code/data/FF_pass_2}\\
-    \textbf{Table 3:} shows the results for the Second pass of the twp pass sort.
+    \textbf{Table 3:} shows the results for the Second pass of the twp pass sort with the EMP factor.
 On the left labeled as (1) are the regression results for the size decile portfolios.
@@
     \paperspacingnarrow
+
     \input{Code/data/All_Industry_FF_pass_2}\\
-    \textbf{Table 5:} SIC top 10\% minus bottom 10\% \hl{NEED TO EDIT THESE RESULTS}
+    \textbf{Table 5:} Shows the results for the Second pass of the twp pass sort with the ESG factor generated from the 10\% of MSCI ESG scores minus bottom 10\% MSCI ESG scores.
+On the left labeled as (1) are the regression results for the size decile portfolios.
+On the right labeled as (2) are the regression results for the beta decile portfolios
+
     \paperspacingwide

### 2021-05-03 Paper/robustness.tex

@@
 To address this robustness concern, I have repeated the two-pass sort with the size decile portfolio by adding a time-based dummy variable.
-Table 7 details the results from the regressions.
+Table 7 details the results from this regression.
 In 2015 2016, 2018, and 2019 there is statistical significance for the EMP term and a positive association between EMP and excess returns.
@@
 In 2019 a one-unit change in beta would result in a 25.9\% increase in excess returns.
-The coefficient values in 2017 appeared to hold a negative relationship as a one-unit change in ESG beta would have resulted in a reduction in excess returns by 10.5\%.
+The coefficient value of the EMP term in 2017  holds a negative relationship as a one-unit change in ESG beta resulted in a reduction in excess returns by 10.5\%.
 Although in 2020, there was a positive relationship, the coefficient is not statistically significant.
@@
 \end{center}
- \textbf{MSCI data 2018-2020}\\
+
+ Moving on to the second set of analyses with a slightly different definition of the ESG factor.
+As described in the mythology and the results, I created another definition for measuring the effect of ESG by subtracting the Best 10\% of frim from the bottom 10\% of firms as defined by MSCI ESG scores.
+As demonstrated in the results section, there was no measured statistical relationship between this factor and the excess returns.
+To further break down this relationship, I decomposed the excess returns on a first digit SIC level to assess if some industries are more sensitive to ESG factors than others.
+Some industries are likely to be inherently ESG unfriendly and might not have a substantial effect from an ESG factor.
+To create this analysis, I made a unique ESG factor for each of the SIC first digit industry codes by subtracting returns from the best 10\% of the firm's ESG score from the bottom 10\% of the firm's ESG scores.
+This process is also detailed in the data section.
+With the new ESG factor, I regressed a two-pass sort using size decile portfolios on each SIC industry to access abnormal returns.
+MSCI did not have data on firms that started with SIC codes 0 or  9.
+Additionally, the MSCI data for SIC code 5 was only BBB; as such, there was no delineation in ESG rating, and I was unable to create a high low ESG factor.
+
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/Industry_FF_pass_2}\\
+    \textbf{Table 8:} shows the reuslts for the second regression reuslts in a two pass sort.
+Columns (1)-(7) represent the first digit SIC code insudtries and their respective sensitivity to the fama french five factor as well as the ESG factor.
+factor\\
+
+    \paperspacingwide
+\end{center}
+
+ Another concern for the MSCI ESG generated ESG factor is the timeframe of the data.stickness to the ESG scores as it takes time to implment changes 
  \begin{center}
@@
     \textbf{Table 8:} SIC top 10\% minus bottom 10\% 2018-2020
+
     \paperspacingwide
 \end{center}
- \textbf{MSCI SIC level analysis}\\
-
-  \begin{center}
-    \paperspacingnarrow
-    \input{Code/data/Industry_FF_pass_2}\\
-    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
-    \paperspacingwide
-\end{center}
-
- WE ADJUSTED THE YEAR to 2019!!
- problems we can never solve for both of these indexes does not do a great job capturing small companies so size might be a bit of a probelm overall as the small companies do not have ther ESG scores baked in as MSCI does not have the resources to rank every company
-
- when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies
-
-dicuss limitation of only one set of ESG factors for MSCI
-
- we used the Size deciles on a go froward basis when analysis the data as the statistical relationship was more apparent

### 2021-05-03 Paper/conclusion.tex
Updates from Overleaf

@@
+The main conclusions from this paper comes from defining and  implementing an ESG factor.
+In the first set of analysis where ESG term is defined as the different between the MSCI ESG index and petroleum index I observed a statistically significant result for the ESG term which matched previous work which has been conducted on ESG factor modeling.
+However in the second set of analysis when developing an ESG factor from the MSCI ratings I was unable to find a statistically significant factor on the aggregate level and I was only able to find a statistically significant factor for \hl{ incert number of industries}.
+This lack of uniformity across both definitions fosters a number of important questions to address in the implementation of FFFF models
+
 Looking beyond this paper's scope and future research opportunities, there are a few natural extensions for assessing the role rating agencies and indexes play in the EGS landscape.

### 2021-05-05 Paper/Results.tex

--- 2021-05-03 Paper/Results.tex
@@
     \input{Code/data/size_output_inital}\\
-    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size.
-the row axis is portfolio and the column axis is the  market beta,four Fama French factors and the EMP factor\\
+    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size deciles.
+The row axis is portfolio and the column axis is the  market beta,four Fama French factors and the EMP factor\\
     \paperspacingwide
 \end{center}
-
 Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for Size decile portfolios.
@@
     \input{Code/data/FF_pass_2}\\
-    \textbf{Table 3:} shows the results for the Second pass of the twp pass sort.
+    \textbf{Table 3:} shows the results for the second pass of the top pass sort with the EMP factor.
 On the left labeled as (1) are the regression results for the size decile portfolios.
@@
     \input{Code/data/beta_output_inital}\\
-    \textbf{Table 4:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
+    \textbf{Table 4:} Shows the first pass in a two pass sort.
+The porflios are decomposed by beta deciles and the table demonstrates the sensitivity of the four factors, market returns and and EMP factor to the decile\\
     \paperspacingwide
@@
     \paperspacingnarrow
+
     \input{Code/data/All_Industry_FF_pass_2}\\
-    \textbf{Table 5:} SIC top 10\% minus bottom 10\% \hl{NEED TO EDIT THESE RESULTS}
+    \textbf{Table 5:} Shows the results for the Second pass of the two pass sort with the ESG factor generated from the top 10\% of MSCI ESG scores minus bottom 10\% MSCI ESG scores.
+On the left labeled as (1) are the regression results for the size decile portfolios.
+On the right labeled as (2) are the regression results for the beta decile portfolios
+
     \paperspacingwide

### 2021-05-05 Paper/motivation.tex

@@
-Asset pricing has a long history of models and theories governing how investors make decisions.
 Over the last decade, investors have become increasingly interested in factors outside of traditional financial metrics when assessing the possibility of an investment's returns.
@@
 
-Global climate change presents a host of possible and sizable financial risks from increased water levels, which could threaten major global cities' safety, including Miami, Honk Kong, and Shanghai, with the potential to displace 100's of millions of people \cite{holder_three-degree_nodate}.
+Global climate change presents a host of possible and sizable financial risks from increased water levels, which could threaten major global cities' safety, including Miami, Hong Kong, and Shanghai, with the potential to displace 100's of millions of people \cite{holder_three-degree_nodate}.
 Global climate change is also shifting the landscape for global agricultural production as more extreme weather conditions, including droughts, pose problems to the most fundamental building block of economic activity.
-In addition to risk mitigation, investors look to utilize ESG principle for superior returns.
-A research report published by McKinsey and Company linked gender diversity on executive teams to higher profitability and superior value creation \cite{mckinsey_diversity}.
-
-In addition to risk investing professionals, ESG investing is appealing to a younger set of investors interested in a plethora of global issues from global climate change to diversity and inclusion in the workplace.
-A study conducted by Morgan Stanley found that millennials investors were twice as likely as the general investing population to invest in a company with social or environmental goals.
+In addition to risk mitigation, investors look to utilize  ESG principles for superior returns.
+A research report published by McKinsey and Company linked gender diversity on executive teams to higher profitability, and superior value creation \cite{mckinsey_diversity}.
 
@@
 
-To address the impact of ESG factors on stock preformance, this paper is broken into five sections.
+In this paper, I assess the impact of ESG investing on the excess market through a five-factor Fama McBeth framework.
+I establish two different methods for assessing ESG risk premia and find dissimilar results.
+When defining ESG risk premia as the difference between the MSCI ESG index and the S\&P Petroleum index, I find a robust risk premia, which suggests that a one-unit change in beta results in an 8.7\% increase in excess returns for size decile portfolios.
+However, when defining ESG risk premia as the difference between high ESG firms minus low ESG firms as rated by MSCI, there are no statistically measurable risk premia.
+Further, when analyzing the risk premia at the one-digit SIC level, there is still no statistical relationship between SIC industry and excess returns and the ESG factor.
+
+I find that the definition of ESG in market return results in vastly different outcomes. When building a market index 
+this paper is broken into five sections.
 \textit{Section \ref{Literature}} will provide a  brief literature review of papers looking at abnormal performance around ESG criteria as well as discussion of asset pricing models to capture ESG factors in performance.

### 2021-05-05 Paper/title_page.tex
big push in latex

@@
 \HRule \\[0.4cm]
-{ \huge \bfseries Factor Investing with EGS}\\[0.4cm] % Title of your document
+{ \huge \bfseries Factor Investing with ESG}\\[0.4cm] % Title of your document
 \HRule \\[1.5cm]
@@
 \begin{flushleft} \large
-\emph{Author:}
-Matthew \textsc{Chistolini}\\ % Your name
+Author: \emph{}
+Matthew Chistolini \textsc{}\\ % Your name
 \end{flushleft}
@@
 % If you don't want a supervisor, uncomment the two lines below and remove the section above
-\Large \emph{Advisor:}
-Michael \textsc{Connolly}\\[2cm] % Your name
+ Advisor: \emph{}
+Michael Connolly \textsc{}\\[2cm] % Your name
 

### 2021-05-05 Paper/robustness.tex
Updates from Overleaf

@@
     \input{Code/data/FF_pass_2_no_ESG}\\
-    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
+    \textbf{Table 6:} Shows  second pass results in a teo pass sort for the size and beta decile portfolios without EMP factor\\
     \paperspacingwide
@@
     \input{Code/data/Time_FF_pass_2}\\
-    \textbf{Table 7:} Shows the results for the second pass in a two pass sort. The two pass sort used size based deciles for the first pass. Columns (1)-(6) represent individual years of data. 
+    \textbf{Table 7:} Shows the results for the second pass in a two pass sort.
+The two pass sort used size based deciles for the first pass.
+Columns (1)-(6) represent individual years of data from 2016-2020.
+
     \paperspacingwide
@@
 Additionally, the MSCI data for SIC code 5 was only BBB; as such, there was no delineation in ESG rating, and I was unable to create a high low ESG factor.
-
+Table 8 demonstrates the results from this new set of regressions.
+For all of the available level SIC portfolios, the market term and the SMB term are highly statistically significant at the 1\% level, which is consistent with the results from both the EMP and the initial MSCI ESG rating analysis.
+The ESG factor in each of the SIC regressions shows a positive but statistically insignificant coefficient even at the 10\% level.
+This result suggests that the EGS factor for the MSCI high low ESG factor is priced into the market, unlike the EMP factor from the first part of the analysis.
 \begin{center}
@@
     \textbf{Table 8:} shows the reuslts for the second regression reuslts in a two pass sort.
-Columns (1)-(7) represent the first digit SIC code insudtries and their respective sensitivity to the fama french five factor as well as the ESG factor.
-factor\\
+Columns (1)-(7) represent the first digit SIC code industries and their respective sensitivity to the fama french five factor as well as the ESG factor.
+SIC code 1 covers the mining and construction industry, SIC codes 2-3 covers manufacturing industry, SIC code 4 covers Transportation, Communications, Electric, Gas and Sanitary service SIC code 6 covers Finance, Insurance, And Real Estate SIC code 7-8 covers services\\
 
@@
 
- Another concern for the MSCI ESG generated ESG factor is the timeframe of the data.stickness to the ESG scores as it takes time to implment changes 
+A final robustness check is conducted to address another concern about the MSCI ESG data.
+As explained in the dad section, the ESG data from MSCI to create the ESG high low index is static.
+That is to say, the data for ESG scores is only a snapshot of ESG ratings for April 2021.
+The use of the entire window from June 2011 to March 2021 to assess excess returns might be a bad characterization of the ESG landscape as a company that could have been a low ESG company might have switched to a high ESG company time.
+To address this concern, I recreated the regression from table 5 by limiting the time horizon to 2018-2020.
+Although there could have been some movement in the ESG scores during the modified time frame, a companies ability to respond to change has a degree of stickiness as corporate initiatives often take several years to develop, and therefore some of the ESG factors would have a lag effect.
+Looking at the results from table 9, the Market SMB and CMA are statistically significant, which was also observed in table 5.
+The ESG factor in table 9, similar to table 5, is not statistically significant for either the beta or size decile portfolios.
+The magnitude of the  ESG factor for the size-based portfolio has increased from .004 to .013; however, the beta decile portfolio has changed signs from .0145 to -.013.
+The muddled resulted mixed with the statistical insignificance do not provide evidence of a shorter time horizon improving the accuracy of the ESG factor as a predictor of stock performance.
+An additional source of error could be a result of the sample bias associated with the companies MSCI has rated.
+MSIC has only rated 545 of the US public equities.
+Of the rated companies 439 (or99\%) have a market cap greater than 1B.As a result of the large market cap bias, the ESG factor does not accurately describe all of the ESG risk associated with the market.
+
  \begin{center}
@@
     \input{Code/data/All_Industry_FF_pass_2_past_2018}\\
-    \textbf{Table 8:} SIC top 10\% minus bottom 10\% 2018-2020
+    \textbf{Table 9:} Shows the same regression from table 5 with an altered time frame to only include data from 2018-2020
 

### 2021-05-07 Paper/Methodology.tex
latex now basically to the end!!
the final push :)

@@
-As discussed in the literature review, there are numerous models for assessing asset pricing and abnormal returns.
+As discussed in the literature review, there are numerous models for assessing asset pricing and excess returns.
 In this paper, I will be using an extension of Fama French's 2015 paper ``A five-factor asset pricing model".
 This analysis method is built on the Fama French three-factor model and relies on the two-pass sort.
-Like the section \ref{Data} on data, this section is broken up into multiple parts.
 First, I will explain the methodology for assessing the sensitivity and risk premium for the ESG factor created from the MSCI ESG index minus the petroleum index.
@@
 The size and beta deciles ratings are provided by CRSP and are merged on the individual firm level for each month of data.
-Next, I conducted the first pass for the two-pass sort by running the regression of expected returns on market risk premia the four Fama French factors and the EMP factor as described in equation \eqref{six_factor_basic_reg}.
+Next, I conducted the first pass for the two-pass sort by running the regression of excess returns on market risk premia, the four Fama French factors, and the EMP factor as described in equation \eqref{six_factor_basic_reg}.
 This regression was run twice: once on the beta-ranked portfolios and again on the size-ranked portfolios.
-
+The first pass provides insight into the sensitivity of the factor on the declie index.
 \begin{equation}
@@
 \end{equation}
-
 \begin{center}
@@
 \end{center}
-
-Next, I used the data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort.
-The second pass of the two-pass sort is described in the equation  \eqref{second_pass}.
-The second pass describes the risk premia for each of the six factors in the mode.
+Next, I used the output data from equation \eqref{six_factor_basic_reg} to run the second regression in the two-pass sort as described in the equation  \eqref{second_pass}.
+The second pass represents the risk premia for each of the six factors in the model.
 This second pass was performed for both beta and size-based portfolios.
-
 \begin{equation}
@@
 \end{equation}\\
-
-The second set of regressions is focused on establishing a link between MSCI ESG ratings and ESG risk factor sensitivity.
+The second set of regressions is focused on establishing a link between MSCI ESG ratings and excess returns.
 Similar to the first set of regressions, I use a two-pass sort methodology using both size and beta ranked portfolios.
-To create the ESG factor for this set of analyses, I used the all industry ESG factor described in Section \ref{Data}.
-The process of implementing the regression follows the same basic form as the one for the MSCI ESG index and petroleum index ESG factor by first running a pass on the portfolio by size and following that with a cross-sectional analysis to quantify the risk premium.
+To create the ESG factor for this set of analyses, I used the ESG factor described in Section \ref{Data}.
+The process of implementing the regression follows the same basic form as the one for the MSCI ESG index and petroleum index ESG factor by first running a pass on the portfolio by size(beta) decile and following that with a cross-sectional analysis to quantify the risk premium.
 
@@
 The newly west error is a replacement for standard error in a traditional OLS regression.
-the advantage of the newly west is an ability to account for serial correlation within the model
+The advantage of the newly west is an ability to account for serial correlation within the model.
+To view any of the supporting code for the analysis, view the link in the citation \cite{Matt_code}.

### 2021-05-07 Paper/Results.tex
the final push :)

--- 2021-05-03 Paper/Results.tex
@@
 The row axis is the distinct size deciles, where 1 is the smallest 10\% of companies by market cap companies, and 10 is the decile with the largest 10\% of companies by market cap.
-The column axis includes estimated coefficients for the intercept, five Fama French factors, and the EMP factor.
+The column axis includes estimated coefficients for the five Fama French factors and the EMP factor.
 $HML_{t}$ is the difference between the returns on diversified portfolios of the high and low book to market stocks.
@@
 $EMP_t$ is the difference between MSCI World ESG Focus Index returns and the S\&P petroleum index.
-Finally, the intercept is the excess returns for the given decile.
 The results in this table demonstrate that for HML, CMA, and EMP factors, there is a size-based relationship wherein the larger the size decile, the smaller the sensitivity effect is on the factor.
-In addition, the EMP factor also exhibits a consistent negative relationship between the EGS factor and returns.
-This relationship indicates that investors are returning a premium by investing in ESG companies over the contra ESG counterparts and that this relationship is largest for small companies.
+This relationship indicates that investors are returning a premium by investing in ESG companies over the contra ESG counterparts and that this relationship is strongest for small companies.
+The size-based sensitivity that is greatest for small firms was also noted in Dorfleitner et al's work\cite{Dorfleitner2020ESGCA}.
 The results from this statistical test are significant at the 5\% level for the EMP factor for all of the portfolios except the smallest and largest size portfolios.
-However, several of the Fama French factors are also not significant for the largest and smallest portfolios, which may indicate that these tail portfolios might be behaving differently due to their composition rather than a reflection of the factor.
+However, several of the Fama French factors are also not significant for the largest and smallest portfolios.
+This finding indicate that these tail portfolios  behave differently and may require winsorizing rather than a reflection of the factor's significance.
 
@@
     \input{Code/data/size_output_inital}\\
-    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size.
-the row axis is portfolio and the column axis is the  market beta,four Fama French factors and the EMP factor\\
+    \textbf{Table 2:} represents the first pass of the two pass sort where the portfolios are broken up by size deciles.
+The row axis is portfolio and the column axis is the  market beta, four Fama French factors, and the EMP factor.\\
     \paperspacingwide
 \end{center}
-
-Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for Size decile portfolios.
+Table 3 demonstrates the risk premia for each of the factors using the second step in the two-pass sort for size decile portfolios.
 The market factor, the SMB factor, the CMA factor, and the EMP factor are all positive and statistically significant at the 1\% level.
 The other two factors have a negative sign and are not statistically significant at the 10\% level.
-EMP factor demonstrates that a 1 unit change in beta corresponds with an 8.7\% increase in returns.
-Although this is by far the smallest coefficient of the four statistically significant returns, this term is still economically meaningful and represents a risk premia for investing in high ESG firms.
-Overall the model appears to be a fairly good fit with a $R^2$ value of 87.5\%, which is lower than some previous literature using the petroleum index but still indicates a well fit model \hl{cite this paper}
+EMP factor demonstrates that a 1 unit change in beta corresponds with an 8.7\% increase in excess returns.
+Although this is by far the smallest coefficient of the four statistically significant factors, this term is still economically meaningful and represents a risk premium can be achieved by investing in high ESG firms.
+Overall the model appears to fit the data reasonably well with a $R^2$ value of 87.5\%, which is lower than some previous literature using the petroleum index but still indicates a well-fit model \cite{Gregory2020TheGP}.
 
@@
     \input{Code/data/FF_pass_2}\\
-    \textbf{Table 3:} shows the results for the Second pass of the twp pass sort.
+    \textbf{Table 3:} shows the results for the second pass of the two pass sort with the EMP factor.
 On the left labeled as (1) are the regression results for the size decile portfolios.
-On the right labeled as (2) are the regression results for the beta decile portfolios\\
+On the right labeled as (2) are the regression results for the beta decile portfolios.\\
     \paperspacingwide
@@
 Rather the relationship appears to have a dome shape where portfolio 6 has the smallest positive sensitivity to the EMP factor, and larger or smaller beta portfolios appear to have a larger sensitivity to the  EMP factor.
-\hl{why might this be}.
-Similar to the size decile analysis, the EMP coefficients are all very significant, with every portfolio having significance at the 5\% level.
+Similar to the size decile analysis, the EMP coefficients are all highly significant, with every portfolio having significance at the 5\% level.
 
-When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB, CMA, and EMP exhibit positive risk premia.
-The other two factors remain statistically insignificant, and the MHL factor changed signs and now has a positive relationship.
-The EMP coefficient indicates that a one-unit change in beta would result in a 7.36\% change in returns for a high ESG company.
-This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude.
-The model is also a good fit with an $R^2$ value of 86.0\%.
-With the baseline analysis proving that there is a relationship between ESG and excess marker returns, I will provide an investigation of the robustness in section \ref{Robustness} by first comparing the results of the two-pass sort with the EGS factor to the same analysis without the factor.
-Next, I will investigate time-related relationships between ESG and excess returns by implementing a dummy variable for the year and repeating the two-pass sort methodology.
+When considering the second phase of the two-pass sort for the beta decile portfolios, the similarities to the size decile continue wherein market returns, SMB, CMA, and EMP exhibit positive risk premia and the other two factors remain statistically insignificant.
+The EMP coefficient indicates that a one-unit change in beta would result in a 7.36\% change in excess returns for a high ESG company.
+This result is slightly smaller than the result from the size portfolio but still demonstrates a statistical relationship in the same order of magnitude and suggests the EGS factor can provide excess returns.
+The beta decile model is a strong fit with an $R^2$ value of 86.0\%.
+
+The two-pass sort indicates that there is a measurable relationship between the EMP factor and excess marker returns.
+In section \ref{Robustness} I will provide an analysis of the robustness by first comparing the results of the two-pass sort with the EGS factor to the same analysis without the factor.
+Next, I will investigate the time-related relationships between ESG and excess returns by implementing a dummy variable for the year and repeating the two-pass sort methodology.
 
@@
     \input{Code/data/beta_output_inital}\\
-    \textbf{Table 4:} Beta decile by estimated coefficients for the  five fama french factors and EMP factor\\
+    \textbf{Table 4:} shows the first pass in a two pass sort.
+The portfolios are decomposed by beta deciles and the table demonstrates the sensitivity of the four factors, market returns and and EMP factor to the decile.\\
     \paperspacingwide
@@
 I repeated the methodology above for the regression analysis.
-I  regressed two sets of two-pass regression analyses, one on beta-sized portfolios and another on size-based portfolios.
-In the first pass results for the MSCI factor, the beta portfolio demonstrated a decrease in sensitivity as the betas grew; however, the size-based portfolios did not experience any real pattern governing the decile and the sensitivity.
-In the case of the size-based portfolios, some of the sensitivity figures were positive while others were negative.
-Table 5 represents a second stage two-pass sort regression output.
-Similar to the regression in the first stage of the analysis, the market return, CMA, SMB remain statistically significant; however, only the size-based portfolios demonstrate a statistically significant coefficient for the ESG factor.
-The EMP ESG factor coefficient can be interpreted as a negative relationship between ESG and excess return wherein a one beta unit increase in ESG results in a decrease of 2\% in excess returns.
-One limitation of the MSCI data is it only represents a snapshot in ESG rating data; that is, the MSCI data is only the most recent ESG performance and does not account for changes in ESG performance.
-As such, a company that used to be a  poor ESG and has improved to a strong ESG company will only be categorized as a strong ESG company which will not accurately address the ESG performance.
-As such, in section \ref{Robustness} I will address this by shrinking the time frame to see if the coefficients increase in their significance.
-Further,  to address additional concerns of robustness, I will run another two-pass sort to isolate the SIC industry-specific ESG factor and determine if there are ESG related excess returns that are greater in some industries than others.
+I  regressed two sets of two-pass sort  regression analyses, one on beta-sized portfolios and another on size-based portfolios.
+In the first pass results for the MSCI ESG factor, the beta portfolio demonstrated a decrease in sensitivity as the beta deciles increased.
+However, the size-based portfolios did not experience any pattern, with some of the coefficients were positive while others were negative.
+
+Table 5 represents the second stage in two-pass sort regression output.
+Similar to the regression with the EMP factor, the market return, CMA, SMB remain statistically significant.
+However, neither the size based size-based or beta-based portfolios demonstrate a statistically significant coefficient for the ESG factor.
+This lack of statistical significance indicates that the MSIC ESG factor is priced into one of the other statistically significant factors and should not produce any excess returns.
+Further solidifying that the factor is not economically significant is the high $R^2$  value for both of the portfolios, which indicates the strong fit of the model.
+
+One potential reason that there is no statistical significance for the factor is the limitation of the MSCI data.
+MSCI only provides current  MSCI ESG rating data.
+As such, a company that used to be a  poor ESG firm that has improved to a strong ESG company will only be categorized as a strong ESG company which will not accurately address the ESG performance.
 
@@
     \paperspacingnarrow
+
     \input{Code/data/All_Industry_FF_pass_2}\\
-    \textbf{Table 5:} SIC top 10\% minus bottom 10\% \hl{NEED TO EDIT THESE RESULTS}
+    \textbf{Table 5:} shows the results for the second pass of the two-pass sort with the ESG factor generated from the top 10\% of MSCI ESG scores minus bottom 10\% MSCI ESG scores.
+On the left labeled as (1) are the regression results for the size decile portfolios.
+On the right, labeled as (2) are the regression results for the beta decile portfolios.
+
     \paperspacingwide
 \end{center}
+
+In section \ref{Robustness} I will address the data concerns by reducing the time frame and observing if the coefficients increase in significance.
+Further, to address additional concerns of robustness, I will run another two-pass sort on individual industries to see if there is uniformity to the risk prima response to the MSCI ESG factor.

### 2021-05-07 Paper/abstract.tex
the final push :)

\pagebreak
\hspace{0pt}
\vfill
\begin{center}
\Huge{Abstract}\\
\normalsize
\vspace{2cm}
ESG investing has more than one definition for  ``good" and ``bad" firms. This paper uses the Fama French five-factor model framework to explore two definitions of EGS risk premia and find radically different results.  When defining the ESG factor as the difference between the MSCI ESG index and the S\&P petroleum index, there is robust evidence for an economically and statistically significant result. This paper finds that a one-unit change in beta results in an 8.7\% increase in excess returns for size decile portfolios. However, changing the ESG factor to the difference between high ESG firms minus low ESG firms as rated by MSCI, there are no statistically measurable risk premia. This lack of uniformity underpins the difficulty of setting uniform policy, which can have a meaningful impact on millions of investor's retirement accounts.  
\end{center}
\vfill
\hspace{0pt}
\pagebreak

### 2021-05-07 Paper/conclusion.tex
the final push :)

--- 2021-05-03 Paper/conclusion.tex
@@
-Looking beyond this paper's scope and future research opportunities, there are a few natural extensions for assessing the role rating agencies and indexes play in the EGS landscape.
+The main conclusions from this paper come from defining and implementing an ESG factor.
+In the first set of analyses where the ESG term is defined as the difference between the MSCI ESG index and the petroleum index, I observed a statistically significant result for the ESG term, which matched previous work conducted on ESG factor modeling.
+However, in the second set of analyses, when developing an ESG factor from the MSCI ratings, I did not find a statistically significant factor on the aggregate level.
+Examining the MSCI results at the one-digit SIC level, I was only able to find the firms starting with SIC code  1  to demonstrate a positive and statistically significant factor.
+These findings reaffirm that the petroleum industry may have negative expected returns but does not build a case for ESG providing excess returns for other industries.
+This lack of uniformity across both definitions fosters a number of important questions to address in the implementation of the Fama french six-factor model.
+This uncertainty in the findings of an ESG factor parody the literature review and make it extremely difficult to recommend a pro ESG policy proposal.
+
+Policy for EGS investing over the last year has been a contentious topic of debate.
+In 2020, the Department of Labor ruled that private pension plan providers ``cannot invest in ESG vehicles that sacrificed investment returns or take on additional risk" \cite{beals_trump_nodate}.
+Investment and retirement planning are essential to the wellbeing of everyone in their retirement, and 401(k) plans are a key part of securing the ability to continue a comfortable life into retirement.
+Prioritizing returns over a longer period,  even by a few basis points can have a sizable impact when considering the effect of compounding over a time frame of 40 years.
+When comparing the current policy position to the results from this study, it would seem like a clear solution to uphold the policy laid out by the Department of Labor.
+However, the results suggest that investors would be better off if pension providers actively avoided traditional fossil fuels as the core of energy production and utilization transitions to renewable solutions.
+
+Further, it is important to note that the results from this study are retrospective and do not account for future development.
+Under the Biden administration, there has been a shift in political tone regarding environmental and social concerns that would support corporate ESG initiatives \cite{newburger_heres_2021}.
+As such, many industries might benefit from these new political tailwinds and create industry ESG divergence similarly to those observed in the petroleum industry.
+
+To account for this uncertainty and opportunity,  I would recommend rewording the mandate from the Bureau of Labor to be more lenient and offer investment solutions that incorporate ESG for those customers who have an active interest in financing global environmental and social reform.
+However,  these investment products should also be provided with additional warnings for the inventor and strict regulations regarding performance claims.
+Return performance should demonstrate the excess return or loss associated with the product's historical performance compared to a not ESG counterpart fund.
+
+Looking beyond this paper's scope to future research opportunities, there are a few natural extensions for assessing the role rating agencies and indexes play in the EGS landscape.
 One extension is to further isolate ESG premia by conducting an event study methodology on several well-known ESG indexes, including the MSCI ESG leaders index referenced in this paper.
 Using addition and deletion events from this index can provide a tangible event to test the market reaction to changes in ESG news and assess if there are abnormal returns.
-This methodology for evaluating abnormal returns can also be applied to ESG rating; however, there might be less of an effect if the market was anticipating a change in the rating.
+This methodology for evaluating abnormal returns can also be applied to ESG rating changes; however, there might be less of an effect if the market was anticipating a change in the rating similar to corporate credit ratings.
 As such, the indexing market event might provide a more precise image regarding market tracking behavior around ESG.
@@
 Over the last few years, more companies have entered into the ESG rating space, MSCI and Sustainalytics have held a dominant position in the market, but S\&P, Moody's, and other firms are looking to expand their research efforts around ESG research.
-An interesting extension would be to analyze the risk premia for different firms and see if there an ESG factor where high ESG minus low ESG has a larger risk premium for one rating agency over another rating agency.
+An interesting extension would be to analyze the risk premia for different firms and see if there is an ESG factor where high ESG minus low ESG has a larger risk premium for one rating agency over another rating agency.

### 2021-05-07 Paper/data.tex
Updates from Overleaf
the final push :)

@@
 Data for the daily returns of MSCI World ESG Focus Index and S\&P petroleum index was sourced from CAPIQ Compustat.
-Data on firm-level ESG scores (by ticker) was sourced from MSCI and reflected the most recent company ratings as MSCI does not have the historical ratings relatively available.
-Finally, daily Fama French factors for SMB, CMA, HML, and RMW were sourced from the Kenneth R.
-French - Data Library.
+Data on firm-level ESG scores (by ticker) was sourced from MSCI and reflected the most recent company ratings.
+Finally, monthly Fama French factors for SMB, CMA, HML, and RMW were sourced from the \textit{Kenneth R.
+French Data Library}.
 
-The data and methodology for the paper are broken into two sections; first, a preliminary study is conducted using a broad base ESG factor to confirm a statistically risk sensitivity relationship between ESG and returns.
-Next, an analysis on a SIC level is conducted in an attempt to understand how industry-specific ESG factors affect the risk-return relationship of stock performance.
+The data and methodology for the paper are broken into two sections; first, a preliminary study is conducted using a broad base ESG factor constructed from two market indexes to confirm a statistical relationship between ESG and excess returns.
+Next, another ESG factor is constructed using the MSCI ESG dataset  to isolate additional ESG risk.
 
 In the first stage of the analysis, an ESG factor is created by taking two market indexes and subtracting their daily returns in an attempt to isolate ESG based risk.
-The formula for this factor is represented in Equation \eqref{making_EMP}, and for the remainder of the section, one analysis will be referred to as EMP, which is sort for ESG minus petroleum.
-This method of creating a factor is very similar to Fama and French in their other four factors in the five-factor model.
+The formula for this factor is represented in Equation \eqref{making_EMP}, and for the remainder of this analysis, this ESG factor will be referred to as the EMP factor (ESG minus Petroleum).
+This method of creating a factor is based on the fama french model used by the other four factors in the five-factor model.
 For this analysis, I will be using the MSCI World ESG Focus Index as the high ESG portfolio and the  S\&P petroleum index as the low ESG portfolio.
-The methodology for creating an ESG factor is similar to the work from \hl{put the author's name in here}.
-In \hl{put the author's name in here}'s work, they were able to find a statistical risk relationship between a high ESG portfolio and a petroleum index\cite{Frynas2005TheFD}.
-The petroleum index provides a good contra ESG portfolio as the petroleum industry is in direct opposition to environmental goals that are E components of ESG.
+The choice of the petroleum index was based on Gregory et al's work where the authors demonstrated the sensitivity of the index to contra ESG factors \cite{Frynas2005TheFD}.
+The petroleum index provides a good contra ESG portfolio as the petroleum industry is in direct opposition to environmental goals that are `E' components of ESG.
 A similar analysis could also be run with other market indexes.
 The S\&P Mining index is another popular index that also performs very poorly for environmental and social concerns\cite{Kapelus2002MiningCS}.
-A final index that also could represent a contra ESG portfolio is the  \$VICE ETF.
-This financial product tracks the market for  alcohol, drug, gaming, and betting industry which have all been shown as clear contra ESG industries \hl{inset citation}
+A third index that also could represent a contra ESG portfolio is the  \$VICE ETF.
+This financial product tracks the market for alcohol, drug, gaming, and betting industry which have all been shown as clear contra ESG industries \cite{Grougiou2016CorporateSR}.
+
 \begin{equation}
@@
 June 2011 marks the initiation of the MSCI ESG index; as such, this is the beginning of the sample period for this paper.
-Daily individual stock return information was collected from CRSP and merged with the daily ESG factor returns.
-The return data was collapsed on month to create a monthly return figure by summing all of the daily returns for each given month in the dataset.
-Next, the Fama French monthly factors were merged into the dataset for each individual firm for each month.
-After aggregating, the factor and individual firm return, there were 830,891 observations of monthly return data for 12,194 unique stocks.
+Daily individual stock return information from CRSP was merged with the daily ESG factor returns.
+The resulting merged data was collapsed on months to create a monthly return figure by summing all of the daily returns for each given month in the dataset.
+Next, the Fama French monthly factors were merged onto the dataset for each individual firm for each month.
+After aggregating, the factor and individual firm return, there were 830,889 observations of monthly return data for 12,194 unique stocks.
 The merged and cleaned dataset had an average of 68 months of data for each individual security.
@@
 
-In the second stage of the analysis, I look at how SIC code-specific ESG factors affect the risk-return relationship.
-In this study, I used the MSCI ESG rating by ticker to created SIC-based ESG factors.
-MSCI rate companies on their ESG score AAA-CCC where AAA is the highest ESG rating and CCC is the lowest ESG rating.
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/desc1}\\
+    \textbf{Table 1:} shows summary statistics for the data set
+    \paperspacingwide
+\end{center}
+
+In the second stage of the analysis, the definition of the ESG factor is changed to isolate additional ESG risk.
+In the first part of the analysis, the contra ESG portfolio's main concern centered on environmental risk.
+To increase exposure to social and governance while maintaining exposure to environmental risk, a portfolio is constructed from MSCI ESG ratings.
+MSCI provides public ratings for 546 firms and issues a rating from AAA-CCC where AAA is the highest ESG rating, and CCC is the lowest ESG rating.
 In the sample of MSCI ratings, 16 companies were AAA, 88 companies were AA, 118 companies were A, 170 companies were BBB, 108 companies were BB, 40 companies were B, and 5 companies earned a CCC rating.
-Each of these ratings was assigned a mating numeric score (1-7) and were merged using the CRSP name file to their corresponding kypermno value by matching ticker information.
-Subsequently, the ESG scores were merged into the monthly return dataset described at the end of the first analysis.
-A final process was conducted to make a unique ESG factor for each of the SIC codes.
-The unique ESG factor was created at the first digit SIC level and generated by splitting the SIC level ESG data into deciles and taking the top 10\% and as the high ESG index bottom 10\% as the low ESG and subtracting high ESG minus low ESG as explained in equation \eqref{making_EMP_section}.
-This ESG factor was created for SIC codes beginning with 1-4 and 6-8.
-SIC code.
-SIC code 5 only had BBB ratings, and MSCI did not have significant data on SIC code 0 or 9.
-In addition to creating industry specific ESG factors I also created a all industry ESG factor using the top 10\% and bottom 10\% for all MSCI ESG scores to create an ESG factor.
-\hl{explain how EMP is the same down here for consistency of reuslts}
+Each of these ratings was assigned a matching numeric score (1-7) and was merged using the CRSP name file to their corresponding kypermno value by matching ticker information.
+The data was merged onto the monthly return dataset described at the end of the first analysis.
+To create an ESG factor with the MSCI rating, numeric ratings were split into deciles, and a high minus low  ESG factor was generated by taking the top 10\% of ESG firms and subtracting that from the bottom 10\%  of ESG firms as described in equation \eqref{making_EMP_section}.
+For the remainder of this analysis, the MSCI ESG factor will be referred to as the ESG factor or the MSCI ESG factor.
 
@@
     \label{making_EMP_section}
-    EMP_i = R_{\text{high ESG,i}}- R_{\text{low ESG,i}}
+    EMP = R_{\text{top 10\% ESG Firms}}- R_{\text{Bottom 10\% ESG Firms}}
 \end{equation}
-
-\begin{center}
-    Where i is the first digit SIC code    
-\end{center}

### 2021-05-07 Paper/lit_review.tex
big push in latex
the final push :)

@@
-This section will be broke into two components.
-The first section will adress  theoretical models for asset purchasing and then we will introduce some of the mythology for  EGS factor in the asset pricing framework.\\
+The modern foundation of asset pricing began with the capital asset pricing model based on Sharp's 1964 work which used a regression analysis to relate market returns to individual security returns \cite{Sharpe1964CAPITALAP}.
+The CAPM model was later iterated on by Fama french in 1993 with the 3 factors, which considered size,  book, and the market terms from the CAPM model.
+Fama French found that these three coefficients created a more explanatory model which demonstrated the relationship that value companies (low book to market) and small companies experience larger excess returns \cite{Fama1992TheCO}.
+In addition to adding the two factors, Fama French also incorporated the two-pass sort methodology, which isolates portfolio sensitivity and risk premia in a multi sep regression process \cite{Kleibergen2013UnexplainedFA, Fama1973RiskRA}.
 
- \hl{add a section on the metholoogy for the 2 pass sort use} \cite{Kleibergen2013UnexplainedFA} and \cite{Fama1973RiskRA} \\
-The modern foundation of asset pricing began with the capital asset pricing model based on Sharp's 1964's work which used a regression analysis to relate market returns to individual security returns \cite{Sharpe1964CAPITALAP}.
-The CAPM model was later iterated on by Fama french in 1993 with the 3 factor, which considered several other terms to predict stock returns.
-Fama and French added size and book to market terms to the CAPM model.
-They found that these three coefficients created a more explanatory model which demonstrated the relationship that value companies(low book to market) and small companies experience larger excess returns \cite{Fama1992TheCO}.
-Fama and French again innovated on the three-factor model in 2015 by adding two factors to the regression.
-In this model, Fama and French consider Investment and profitability while holding the other regression stable factors.
+Fama and French innovated on the three-factor model in 2015 by adding two additional regressors, investment and profitability.
 The five-factor model analysis found that the addition of the investment and profitability factor made the value factor redundant \cite{Fama2013AFA}.
-The five-factor framework analysis has been recreated on international exchanges to test the effects on foreign markets.
-In recreation by Fama French and another study by Zhong and Li, the general relationships demonstrated in the five-factor model are also applicable in China, Japan, and Australian markets \cite{Fama2015InternationalTO, LIN2017141, Chiah2016A}.
-However, when the Fama French regression was applied to the Australian equity market, the authors did not find the significance of the value factor to subside like was visible in the 2015 paper \cite{Chiah2016A}.
-In addition to expanding the five-factor analysis geographically, some scholars have expanded the five-factor model by adding additional terms to assess momentum effects.
-In a paper published by  Dirkx and  Peter, the authors expanded the five-factor model by applying the Fama French framework to the German market while also adding a sixth term to assess the effect of risk premium with the CDAX constituents' momentum.
-Another study by Gregory and Stead used the sixth factor as an ESG factor to analyze if the ESG effect played a significant role in determining risk premia.
-This paper found that  "the risk premium of the sustainability factor is positive and significant\cite{Gregory2020TheGP}."
+The reliability of the five-factor has been shown by replications on international exchanges and has been found to hold true in Germany, China, and Japan but did not hold true for Australia \cite{Fama2015InternationalTO, LIN2017141, Chiah2016A}.
+The Fama French five factor model is the standard in asset pricing and is the basis for many academic research projects that extend the five-factor model to address additional unpriced factors \cite{Dirkx2020TheFF}.
+Gregory and Stead use a sixth factor as an ESG factor to analyze if the ESG effect played a significant role in determining risk premia.
+This paper found that  ``the risk premium of the sustainability factor is positive and significant"\cite{Gregory2020TheGP}.
 
+Literature on ESG investing has become a quickly growing academic interest, with researchers finding both strong evidence for a non-priced ESG factor and no evidence of a noticeable ESG factor.
 In addition to Gregory and Stead's paper, several other authors have used the Fama French model to address the question concerning ESG and risk premia.
@@
 With the ranked ESG score data, the author then creates two portfolios by separating the top decile score and the bottom deciles scores, which act as the ESG and contra ESG portfolios.
-Similar to Gregory et al.
-paper, the author subtract the top 10\% from the bottom 10\% to create the ESG factor \cite{Dorfleitner2020ESGCA}.
+The author finds that there are measurable excess returns and contributes.
+The main findings for this paper is the realization that small companies, particularly those who stay out of controversy, trend to be undervalued and appreciate compared to similar size companies with lots of controversy \cite{Dorfleitner2020ESGCA}.
+
+Although authors have found positive evidence, other authors have found no such evidence of a relationship.
+Work completed by Liou investigates the MSCI ESG rating data to construct equal-weighted portfolios as far back as 1991  and performs a Fama French analysis.
+In the paper, the author finds that in a post-GFC test, there is no excess return for a long strong ESG short poor ESG factor.
+However, the author does note that there is a relationship to the size and the coefficient for the ESG factor \cite{Lioui2018ESGFI}.
+
+Additionally, some authors have looked at the excess returns on ESG portfolios and assess the effects over calendar years.
+Work from Bennani et al.
+explored the post-GFC landscape for ESG investing and found that in North American stocks from 2010-2013, investing with ESG screening reduced excess returns; however, from 2014-2017, the authors found that ESG screening created 3.3\% excess returns annually \cite{Bennani2018HowEI}.
+%.
+These results strongly indicate that this is, on the one hand, driven by low-rated smaller companies (“small sinners”) and clean-coated frms with regard to controversies (“silent saints”) on the other hand.
+
+%Among others, Derwall et al.
+(2005) and Edmans (2011), who link the doing good while doing well-hypothesis with the managerial myopia theory, conclude that short-term investors are unable (or unwilling) to price the long-term benefts of those activities correctly and therefore undervalue stocks of companies with high levels of engagement in environmental or social aspects, leading to higher returns in the long-run for the respective stocks when compared with other stocks.
+%with the fndings of Dorfeitner et al.
+(2018), who conclude that the benefts of socially responsible activities (measured by the abnormal stock returns) are produced by unexpected additional cash fows which occur mid-to-long term.
+%Furthermore, from an investor’s standpoint, having a “clean coat” with regard to controversies is especially proftable for smaller companies, as the absence of these scandals may be overlooked and incorrectly incorporated in the market prices.
+Thus, one might say that the respective companies “fy under the radar”.

### 2021-05-07 Paper/motivation.tex
the final push :)

--- 2021-04-29 Paper/motivation.tex
@@
-Asset pricing has a long history of models and theories governing how investors make decisions.
-Over the last decade, investors have become increasingly interested in factors outside of traditional financial metrics when assessing the possibility of an investment's returns.
-Much of this enthusiasm has been based on an ESG, Environmental Social Governance, framework.
-ESG investing considers how firms respond to these factors as potential impetuses for excess performance.
-This increased attention is especially visible in Morgan Stanley's 2019 survey of the ESG landscape.
-Morgan Stanley  found that 75\% of assessment managers say that "their firm has adopted sustainable investing." Only two years prior, in 2016, just 10\% of surveyed asset managed had adopted sustainable investing, demonstrating the rapid speed of adoption for ESG investing analysis \cite{morgan_stanley_sustainable_2019}.
-Practitioners have utilized the ESG framework to reduce risk in their investments while also finding abnormal returns.
+Over the last decade, investors have become increasingly interested in monitoring factors beyond traditional financial metrics to find opportunities to achieve excess returns.
+In this pursuit,  many investors have focused their attention on the ESG, Environmental Social Governance, framework.
+Using the ESG framework, investors evaluate a firm's sustainability and societal impact, which are typically overlooked in traditional asset pricing models.
+This method of investing has garnered quick traction as global climate change threatens the longevity of critical global infustructure, and diversity initiatives shape the modern workforce  \cite{holder_three-degree_nodate, mckinsey_diversity}.
+A study conducted by Morgan Stanley found in 2016, just 10\% of surveyed asset managers had adopted sustainable investing, a number that quickly grew to over 70\% by 2018 \cite{morgan_stanley_sustainable_2019}.
 
-Global climate change presents a host of possible and sizable financial risks from increased water levels, which could threaten major global cities' safety, including Miami, Honk Kong, and Shanghai, with the potential to displace 100's of millions of people \cite{holder_three-degree_nodate}.
-Global climate change is also shifting the landscape for global agricultural production as more extreme weather conditions, including droughts, pose problems to the most fundamental building block of economic activity.
-In addition to risk mitigation, investors look to utilize ESG principle for superior returns.
-A research report published by McKinsey and Company linked gender diversity on executive teams to higher profitability and superior value creation \cite{mckinsey_diversity}.
+This new wave of investor sentiment has created strong tailwinds for sustainable investment products.
+In 2020 alone, fund flows into sustainable funds grew by \$51 billion  which is up over 100\% from the 2019 fund flow into sustainable investment products of about \$21 billion, placing total assets under management in sustainable products at over \$250 billion.
+Fund flows into sustainable products for 2020 accounted for a little less a quarter of all US mutual fund flow and about one-fifteenth of all ETF fund flow according to Morningstar \cite{monring_star_ESG}.
 
-In addition to risk investing professionals, ESG investing is appealing to a younger set of investors interested in a plethora of global issues from global climate change to diversity and inclusion in the workplace.
-A study conducted by Morgan Stanley found that millennials investors were twice as likely as the general investing population to invest in a company with social or environmental goals.
-
-This new wave of investor sentiment has created strong tailwinds for sustainable investment projects.
-In 2020 alone, fund flows into sustainable funds grew by 51 billion dollars which is up over 100\% for the 2019 fund flow into sustainable investment products of about 21, placing total assets under management in sustainable products at over 250B.
-Fund flows into sustainable products for 2020 accounted for a little less a quarter of all US mutual fund flow and about one-fifteenth of all ETF fund flow according to morningstar \cite{monring_star_ESG}.
-
-To address the impact of ESG factors on stock preformance, this paper is broken into five sections.
-\textit{Section \ref{Literature}} will provide a  brief literature review of papers looking at abnormal performance around ESG criteria as well as discussion of asset pricing models to capture ESG factors in performance.
-Next \textit{ Section \ref{Data}}  will provide a brief discussion of the data used in this paper as well as the data cleaning process and summary statistics.
-Following the Data, \textit{Section \ref{Methodology}} will breakdown of the methodology for the paper.
-Finally, \textit{Section \ref{Results}}  will summarize the results of the regressions, \textit{Section \ref{Robustness}} will check the robustness of the results and \textit{Section \ref{Conclusion}} will provide a conclusion and possible extensions for next steps.
+In this paper, I assess the impact of ESG investing on the excess returns through a five-factor Fama French framework.
+I establish two different methods for defining ESG risk premia and find dissimilar results.
+When defining ESG risk premia as the difference between the MSCI ESG index and the S\&P petroleum index, I find a robust risk premia, which suggests that a one-unit change in beta results in an 8.7\% increase in excess returns for size decile portfolios.
+However, when defining ESG risk premia as the difference between high ESG firms minus low ESG firms as rated by MSCI, there are no statistically measurable risk premia.
+Further, when analyzing the risk premia at the one-digit SIC level, only one SIC code yields a statistical relationship between excess returns and the ESG factor.

### 2021-05-07 Paper/robustness.tex
the final push :)

--- 2021-05-03 Paper/robustness.tex
@@
-As discussed in section \ref{Results}, additional tests are required to ensure the methodology's robustness.
-The first robustness check I will run is a repetition of the two-pass sort for where the ESG factor is calculated asMSCI ESG index minus the petroleum index (EMP).
+As discussed in section \ref{Results}, additional tests are required to ensure the robustness of the methodology.
+The first robustness check I will run is a repetition of the two-pass sort for where the ESG factor is calculated as MSCI ESG index minus the petroleum index (EMP).
 Table 6 confirms the statistical significance of the market term, the SMB term, and the CMA term.
-All three of these terms were found to be statistically significant in the initial regression as well.
 Beyond the coefficients,  the results in table 6 demonstrate a slightly lower $R^2$ for both the Beta decile and the size decile variants, demonstrating that the EMP term contributes to the explanation of excess returns.
 In addition, the constant term for both size and beta decile portfolios increased in magnitude as a result of omitting the EMP.
-As the coefficient term gets closer to zero, the factors do a better job describing the returns and indicate that there is no excess return that is not explained.
- %as dicussed in section \ref{Data} [talk about using mining instead of petrolum ]
+As the coefficient term gets closer to zero, the factors do a better job describing the excess return and relationship and eliminates all other sources of unexplained excess returns.
 
@@
     \input{Code/data/FF_pass_2_no_ESG}\\
-    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
+    \textbf{Table 6:} shows  second pass results in a teo pass sort for the size and beta decile portfolios without EMP factor\\
     \paperspacingwide
@@
 Another robustness check that is essential to consider is the time-based effects on the model.
-Although we have demonstrated that the EMP factor is statistically significant and explains part of our model, it is essential to understand if this is ubiquitous across time in the sample or if the  ESG factor has shifted in risk premia over time.
+Although I have demonstrated that the EMP factor is statistically significant and explains part of our model, it is essential to understand if this is ubiquitous across time in the sample or if the  ESG factor has shifted in risk premia over time.
 To address this robustness concern, I have repeated the two-pass sort with the size decile portfolio by adding a time-based dummy variable.
-Table 7 details the results from the regressions.
+Table 7 details the results from this regression.
 In 2015 2016, 2018, and 2019 there is statistical significance for the EMP term and a positive association between EMP and excess returns.
-The ESG factor showed the most strength in 2019 as well as the greatest coefficient value.
+The ESG factor showed the most statistical strength in 2019 with the greatest excess return.
 In 2019 a one-unit change in beta would result in a 25.9\% increase in excess returns.
-The coefficient values in 2017 appeared to hold a negative relationship as a one-unit change in ESG beta would have resulted in a reduction in excess returns by 10.5\%.
-Although in 2020, there was a positive relationship, the coefficient is not statistically significant.
-However, in 2020, the CMA factor, which has been statistically significant at the 1\% level for all other tests, also does not appear to be statistically significant.
-2020 represents a bit of a deviation due to the larger market condition from the COVID-19 pandemic rather than the EMP factor's long-run behavior.
+These results are consistent with Bennani et al.'s paper decomposing risk premia over time for 2015 and 2016.
+The coefficient value of the EMP term in 2017  holds a negative relationship as a one-unit change in ESG beta resulted in a reduction in excess returns by 10.5\%.
+In 2020, there was a positive relationship, however the coefficient is not statistically significant.
+The CMA factor, which has been statistically significant at the 1\% level for all other tests, also does not appear to be statistically significant for 2020.
+The 2020 calendar year represents a bit of a deviation from typical trends likely due to the larger market condition from the COVID-19 pandemic rather than the EMP factor's long-run behavior.
+In addition to this set of robustness checks, it would be beneficial to repeat the series of tests replacing the S\&P petroleum index with the S\&P mining index as the contra ESG portfolio.
+Both of the proposed indexes fall under the first digit  SIC code 1  and are exposed to similar ESG risks; however, the additional test can isolate risk premia and test if the factor is capturing petroleum industry effect or a robust ESG factor.
 
@@
     \input{Code/data/Time_FF_pass_2}\\
-    \textbf{Table 7:} Shows the results for the second pass in a two pass sort. The two pass sort used size based deciles for the first pass. Columns (1)-(6) represent individual years of data. 
+    \textbf{Table 7:} shows the results for the second pass in a two pass sort.
+The two pass sort used size based deciles for the first pass.
+Columns (1)-(6) represent individual years of data from 2015-2020.
     \paperspacingwide
 \end{center}
- \textbf{MSCI data 2018-2020}\\
+
+ Moving on to the second set of robustness checks with a slightly different definition of the ESG factor.
+As described in the methodology and the results, I created another definition for measuring the effect of ESG by subtracting the Best 10\% of frim from the bottom 10\% of firms as defined by MSCI ESG scores.
+As demonstrated in the results section, there was no measured statistical relationship between this factor and the excess returns.
+To further break down this relationship, I decomposed the excess returns on a first digit SIC level to assess if some industries are more sensitive to ESG factors than others.
+Some industries are likely to be inherently ESG unfriendly and might not be substantially affected by an ESG factor.
+
+To create this analysis, I repeated the two-pass sort methodology from the results using the size decile portfolios and isolated the returns for each of the first digit SIC code industries.
+I regressed the same ESG factor described in the results on this subset to assess the different effects the factor has on each industry.
+As the MSCI ESG rating dataset for this analysis did not have robust data for industries with first digit SIC codes 0, 5, or 9, I elected not to analyze these SIC industries as the construction of the factor did not account for them.
+
+Table 8 demonstrates the results from this new set of regressions.
+For all of the available  SIC portfolios, the market term and the SMB term are highly statistically significant at the 1\% level, which is consistent with the results from both the EMP and the initial MSCI ESG rating analysis.
+The ESG factor response to each industry is not uniform.
+Industries with first digit SIC codes 3,4,6,7, and 8 have small non statistically significant coefficients.
+The first digit SIC code 2 demonstrated a statistically significant negative relationship to the ESG factor, demonstrating that the high ESG beta portfolio would harm excess returns.
+However, the first digit SIC code 1 demonstrated a highly statistically significant positive coefficient.
+A one-unit increase in ESG beta would result in an 11.6\% increase in excess returns for first digit SIC industry 1.
+These results confirm results from part 1 of the analyses with the EMP factor as the SIC code 1 encompasses the petroleum industry.
+Future tests should be constructed to ensure that the ESG factor is picking up on more than negative excess returns from investing in the petroleum industry.
+
+\begin{center}
+    \paperspacingnarrow
+    \input{Code/data/NEW_Industry_FF_pass_2}\\
+    \textbf{Table 8:} shows the results for the second regression results in a two-pass sort.
+Columns (1)-(7) represent the first digit SIC code industries and their respective risk premia to the Fama French five-factor as well as the ESG factor.
+SIC code 1 covers the mining and construction industry, SIC codes 2-3 covers manufacturing industry, SIC code 4 covers transportation, communications, electric, gas, and sanitary service SIC code 6 covers finance, insurance, and real estate SIC code 7-8 covers services\\
+
+    \paperspacingwide
+\end{center}
+
+A final robustness check is conducted to address another concern about the MSCI ESG data.
+As explained in the data section, the MSCI ESG data used to create the factor is static.
+That is to say, the data for ESG scores is only a snapshot of ESG ratings for April 2021.
+The use of the entire window from June 2011 to March 2021 to assess excess returns might be a bad characterization of the ESG landscape as a company that could have been a low ESG company might have switched to a high ESG company in that time.
+To address this concern, I recreated the regression from table 5 by limiting the time horizon to 2018-2021.
+Although there could have been some movement in the ESG scores during the modified time frame, a companies ability to respond to ESG objectives has a degree of stickiness as corporate initiatives often take several years to develop and implement, leading to a lag effect.
+Looking at the results from table 9, the market, SMB, and CMA factors are all statistically significant, which was also observed in table 5.
+The ESG factor in table 9, similar to table 5, is not statistically significant for either the beta or size decile portfolios.
+The magnitude of the  ESG factor for the size-based portfolio has increased from .004 to .013, which indicates the EGS factor might have a larger impact on excess returns.
+However, the beta decile portfolio has changed signs from .0145 to -.013, which indicates that the EGS factor has a negative impact on excess returns.
+The muddled resulted mixed with the statistical insignificance do not provide evidence of a shorter time horizon improving the accuracy of the ESG factor as a predictor of stock performance.
+An additional source of error could be a result of the sample bias associated with the companies MSCI has rated.
+MSCI's publicly available ESG ratings data has 545 US firms.
+Of the rated companies, 439 (or 99\% of rated firms) have a market cap greater than \$1 billion.
+As a result of the large market cap bias, the ESG factor might not accurately reflect the risk premia associated with all ESG investing.
+
  \begin{center}
@@
     \input{Code/data/All_Industry_FF_pass_2_past_2018}\\
-\textbf{Table 8:} SIC top 10\% minus bottom 10\% 2018-2020
+    \textbf{Table 9:} shows the same regression from table 5 with an altered time frame to only include data from 2018-2021
     \paperspacingwide
 \end{center}
- \textbf{MSCI SIC level analysis}\\
-
-  \begin{center}
-    \paperspacingnarrow
-    \input{Code/data/Industry_FF_pass_2}\\
-    \textbf{Table 6:} second pass results for the size and beta decile portfolios without ESG factor\\
-    \paperspacingwide
-\end{center}
-
- WE ADJUSTED THE YEAR to 2019!!
- problems we can never solve for both of these indexes does not do a great job capturing small companies so size might be a bit of a probelm overall as the small companies do not have ther ESG scores baked in as MSCI does not have the resources to rank every company
-
- when doing the SIC porlfios did not have data for SIC starting with 5 b/c they were all BB companies
-
-dicuss limitation of only one set of ESG factors for MSCI
-
- we used the Size deciles on a go froward basis when analysis the data as the statistical relationship was more apparent
